---
title: "Classification_results"
author: "Niek Brouwer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Classification results

This document analyses the findings from the logistic regression experiments in the preditions_fig5.Rmd file.
We proceed with the following steps:
1. Identify features that characterize and separate PAM50 subtypes.
2. Visualize and explain feature selection.
3. Conclude: is density + distance enough? Hypothetical additonal method measuring network-based or higher-order sturctures.
4. Identify features that find subclusters in PAM50 subtypes.
5. Measure association of subclusters with clinical features & structure presence.

```{r}
library(here)
library(ggplot2)
library(tidyverse)
library(fst)
source(here("UtilityFunctions.R"))
source(here("MEP_UtilityFunctions.R"))
library(ggpubr)
library(rstatix)
library(ComplexHeatmap)
library(factoextra)

```

```{r}
clinical_data <- getClinical()
density_features <- getDensityFeatures()
shape_features <- getShapeFeatures()
scale_features <- getScaleFeatures()

shape_featuresScaled <- generate_matrix(shape_features, 'tnumber')
scale_featuresScaled <- generate_matrix(scale_features, 'tnumber')
```


## Step 1 & 2
```{r}
generate_boxplot <- function(subtype,df, feature){
    selected <- merge(df %>% select(all_of(c("ImageNumber", feature))), clinical_data %>% select(c('ImageNumber', 'PAM50')), by='ImageNumber', all.x=T) %>% rename('f' = feature)
      stat.test <- selected %>%
     t_test(f ~ PAM50, paired = FALSE, p.adjust.method = "bonf", ref.group = subtype)
  stat.test <- stat.test %>% add_xy_position(x = 'PAM50')
  stat.test$custom.label <- ifelse(stat.test$p.adj <= 0.05, stat.test$p.adj, "ns")

  p <- ggboxplot(selected, x='PAM50', y='f') + theme_bw()  +
    stat_pvalue_manual(stat.test, label = "custom.label") +
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.10))) + ylab(feature)

  return(p)
}
```

```{r}
generate_heatmap <- function(subtype, df, features, column_order=T){
    selected <- df %>% select(any_of(features))
    subtypes <- merge(tibble('ImageNumber' = rownames(df)), clinical_data %>% select(c('ImageNumber', 'PAM50')), by='ImageNumber', all.x=T) %>% mutate(PAM50 = ifelse(PAM50 == subtype, subtype, 'other')) %>% pull(PAM50)
    
    # row_ha = rowAnnotation(subtype = subtypes)

    p <- Heatmap(as.matrix(selected), cluster_columns =  column_order,row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=8), row_split = subtypes)
    return(p)
}
```

```{r}
generate_labels <- function(matrix){
labels <- merge(as_tibble(rownames(matrix)), clinical_data %>% select(c('ImageNumber', 'PAM50')), by.x='value', by.y='ImageNumber', all.x=T)
labels <- labels %>% rename(ImageNumber = value)%>%
  mutate(HER = ifelse(PAM50=='HER2', 'HER2', 'Other')) %>%
  mutate(LuminalA = ifelse(PAM50=='Luminal A', 'Luminal A', 'Other')) %>%
  mutate(LuminalB = ifelse(PAM50=='Luminal B', 'Luminal B', 'Other')) %>%
  mutate(Bas = ifelse(PAM50=='Basal', 'Basal', 'Other')) %>%
  mutate(Norm= ifelse(PAM50=='Normal-like', 'Normal', 'Other'))

colnames(labels) <- gsub(" ", "_", colnames(labels))
return(labels)
}

plot_pca <- function(matrix, label){
  labels <- generate_labels(matrix)
  matrix_withlabels <- cbind.data.frame(matrix, rownames(matrix))
  colnames(matrix_withlabels)[ncol(matrix_withlabels)] = 'ImageNumber'
  matrix_withlabels <- merge(matrix_withlabels, labels, by='ImageNumber', all.x=T)
  
  res.pca <- prcomp(matrix_withlabels %>% select(-colnames(labels)), scale = FALSE)
  print(fviz_eig(res.pca))
  print(fviz_pca_biplot(res.pca, label="var", habillage=matrix_withlabels %>% pull(label), addEllipses = T, ellipse.level = 0.95,select.var = list(contrib = 5) ))
  
}

pca_on_subset <- function(subtype, features){
  densities <- getDensityFeatures() %>% select(any_of(c('ImageNumber', features[["density"]])))
  shapes <- getShapeFeatures() %>% rename(ImageNumber = tnumber) %>% select(any_of(c('ImageNumber', features[["shape"]])))
  scales <- getScaleFeatures() %>% rename(ImageNumber = tnumber) %>% select(any_of(c('ImageNumber', features[["scale"]])))
  
  allDF <- merge(merge(densities, shapes, by='ImageNumber', all.x=T, all.y=T),scales, by='ImageNumber', all.x=T, all.y=T)

  allDF_matrix <- generate_matrix(allDF, 'ImageNumber')
  
  plot_pca(allDF_matrix, subtype)
  
}
```


### Basal

```{r}
Basal_density_features <- c('Basal', "Ep Ki67^{+}", "MHC^{hi}CD15^{+}","Myofibroblasts", "CD38^{+} lymphocytes")

for (f in Basal_density_features){
  print(generate_boxplot('Basal', density_features, f))
}


```


```{r}
scale_up <- c('Endothelial_to_Myofibroblasts', 'Myofibroblasts_to_Myofibroblasts', 'Basal_to_Myofibroblasts','CK^{med}ER^{lo}_to_Myofibroblasts',	'CK^{med}ER^{lo}_to_Fibroblasts',	'CK8−18^{hi}CXCL12^{hi}_to_CK8−18^{hi}CXCL12^{hi}', 'CK8−18^{hi}CXCL12^{hi}_to_CK^{med}ER^{lo}','CK^{med}ER^{lo}_to_ER^{hi}CXCL12^{+}','Ep Ki67^{+}_to_CK^{med}ER^{lo}',	'CK^{med}ER^{lo}_to_CK8−18^{hi}CXCL12^{hi}',
                                            'CK^{+} CXCL12^{+}_to_CK^{med}ER^{lo}','CK^{+} CXCL12^{+}_to_CK^{lo}ER^{lo}','Myofibroblasts_to_CK^{+} CXCL12^{+}')
  
scale_down <- c('Macrophages_to_Myofibroblasts PDPN^{+}','Myofibroblasts PDPN^{+}_to_Endothelial','Macrophages_to_CK^{lo}ER^{lo}','Macrophages_to_Macrophages', 'CD8^{+} T cells_to_Myofibroblasts','CK^{lo}ER^{lo}_to_CK^{lo}ER^{lo}','CK^{med}ER^{lo}_to_CK^{lo}ER^{lo}','CK8−18^{hi}CXCL12^{hi}_to_Myofibroblasts', 'Endothelial_to_CK8−18^{hi}CXCL12^{hi}')

features <- c(scale_up, scale_down)

p1 <- generate_heatmap('Basal',scale_featuresScaled, features, column_order = T )
save_pdf(p1, filename = here('output/Method_comparison/feature_selection/basal_scale.pdf'))

```

```{r}
Basal_density_features <- c('Basal',  "CD38^{+} lymphocytes")
features <- list(density = Basal_density_features, shape = c(), scale = c(scale_up))
pca_on_subset('Bas', features)

```


### HER2

```{r}
generate_boxplot('HER2', density_features,  "CD15^{+}" )
generate_boxplot('HER2', density_features,  "HER2^{+}" )

```
```{r}
shape_up <- c('Fibroblasts_to_CK^{+} CXCL12^{+}', 'CK^{lo}ER^{lo}_to_Endothelial','Myofibroblasts_to_CK^{+} CXCL12^{+}', 'CK^{med}ER^{lo}_to_HER2^{+}')
shape_down <- c('Fibroblasts_to_Endothelial', 'Endothelial_to_CK^{med}ER^{lo}','CK^{med}ER^{lo}_to_CK^{lo}ER^{lo}','Myofibroblasts_to_CK^{med}ER^{lo}','Fibroblasts_to_Myofibroblasts PDPN^{+}','Myofibroblasts PDPN^{+}_to_CK^{med}ER^{lo}')
scale_up <- c('Macrophages_to_Fibroblasts')

p1 <- generate_heatmap('HER2',shape_featuresScaled, c(shape_up, shape_down), column_order = T )
save_pdf(p1, filename = here('output/Method_comparison/feature_selection/HER2_shape.pdf'))

p1 <- generate_heatmap('HER2',scale_featuresScaled, scale_up, column_order = T )
save_pdf(p1, filename = here('output/Method_comparison/feature_selection/HER2_scale.pdf'))
```


```{r}
features <- list(density = c("HER2^{+}"), shape = c(shape_up, shape_down), scale = c(scale_up))
pca_on_subset('HER', features)
```



### Luminal A
```{r}
LumA_density_features <- c("CK^{+} CXCL12^{+}","CK^{lo}ER^{lo}","CK8-18^{+} ER^{hi}","CK8-18^{hi}CXCL12^{hi}","CK8-18^{hi}ER^{lo}","ER^{hi}CXCL12^{+}", "Endothelial","Fibroblasts"  )

for (f in LumA_density_features){
  generate_boxplot('Luminal A', density_features,  f )
}

```

```{r}
LumA_density_features <- setdiff(LumA_density_features, c("CK^{lo}ER^{lo}","CK8-18^{hi}ER^{lo}","Fibroblasts"))
scale_up <- c('Macrophages_to_Macrophages')

p1 <- generate_heatmap('Luminal A',scale_featuresScaled, scale_up, column_order = T )
save_pdf(p1, filename = here('output/Method_comparison/feature_selection/LuminalA_scale.pdf'))

features <- list(density = LumA_density_features, shape = c(), scale = c(scale_up))
pca_on_subset('PAM50', features)
```


### Luminal B

```{r}
LumB_density_features <- c("CK^{med}ER^{lo}", "CK^{lo}ER^{med}" )

for (f in LumB_density_features){
  print(generate_boxplot('Luminal B', density_features,  f ))
}

```
```{r}
LumB_density_features <- c("CK^{med}ER^{lo}")
scale_up <- c('CK^{lo}ER^{lo}_to_Fibroblasts', 'CK^{lo}ER^{lo}_to_Myofibroblasts' )
scale_down <- c('CK^{med}ER^{lo}_to_CK^{med}ER^{lo}', 'ER^{hi}CXCL12^{+}_to_CK^{med}ER^{lo}')

p1 <- generate_heatmap('Luminal B',scale_featuresScaled, c(scale_up,scale_down), column_order = T )
save_pdf(p1, filename = here('output/Method_comparison/feature_selection/LuminalB_scale.pdf'))

features <- list(density = LumB_density_features, shape = c(), scale = c(scale_down, scale_up))
pca_on_subset('PAM50', features)


```



### Normal-like

```{r}
generate_boxplot('Normal-like', density_features, "B cells" )
generate_boxplot('Normal-like', density_features, "Myofibroblasts PDPN^{+}" )

```
### ALL

```{r}
generate_boxplot <- function(df, feature){
    selected <- merge(df %>% select(all_of(c("ImageNumber", feature))), clinical_data %>% select(c('ImageNumber', 'PAM50')), by='ImageNumber', all.x=T) %>% rename('f' = feature)
    
    compare_means(f ~ PAM50,  data = selected,
              ref.group = ".all.", method = "t.test")
    
    # Visualize the expression profile
    p <- ggboxplot(selected, x = 'PAM50', y = 'f', color = "PAM50",
              add = "jitter", legend = "none") +
      ylab(feature) +
      rotate_x_text(angle = 45)+
      geom_hline(yintercept = mean(selected$f), linetype = 2)+ # Add horizontal line at base mean
      stat_compare_means(method = "anova", label.y = 1600)+        # Add global annova p-value
      stat_compare_means(label = "p.format", method = "t.test",
                         ref.group = ".all.", hide.ns = TRUE)      # Pairwise comparison against all


  return(p)
}

for (f in getTumorAndTMETypes()$tumor_types){
  print(generate_boxplot(density_features,  f ))
}

for (f in getTumorAndTMETypes()$TME_types){
  print(generate_boxplot(density_features,  f ))
}
```

