---
title: "Danenberg Reproduction"
author: "Niek Brouwer"
date: "`r Sys.Date()`"
output: html_document
---

```{r introduction, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we aim to reproduce the results of the paper Breast tumor microenvironment structures are associated with genomic features and clinical outcome by Danenberg et al. (https://www.nature.com/articles/s41588-022-01041-y).
We focus on the description of the spatial organization of the TME as presented in figures 4b,d,f and its prognostic impact as seen in figure 6.

The separation and phenotype classification of single cells is not reproduced and this is retrieved from the original paper.

```{r libraries and settings, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(data.table)
library(parallel)
library(igraph)
library(here)
library(fst)

source(here("UtilityFunctions.R"))

cells <- read_fst(here("DATA/SingleCells.fst"), as.data.table = T)
cellNeighbours <- read_fst(here("DATA/CellNeighbours.fst"), as.data.table = T)
n_threads <- parallel::detectCores()
```

## Community detection
1. Create spatial graphs and 2. detect communities based on random walks

```{r community detection}

getCommunities <- function(dt){
	g <- graph_from_data_frame(dt, directed = F)
	subgraphs <- cluster_walktrap(g) #Community-detection algorithm
	assignments <- membership(subgraphs)
	matchOrder <- match(dt[[1]], as.numeric(names(assignments)))	
	communities <- assignments[matchOrder]
	return(data.table(communities))
}

getNetworkProperties <- function(dt){
	g <- graph_from_data_frame(dt, directed = F)
	# n of edges
	nEdges <- ecount(g)
	# n of vertices
	nVertices <- vcount(g)
	# length of the longest path (in number of edges) between two nodes
	diameter <- diameter(g) 
	# average number of edges between any two nodes in the network.
	distance <- mean_distance(g)
	# ratio of the number of edges and the number of possible edges
	density <- edge_density(g)
	# probability that the adjacent vertices of a vertex are connected. 
	transitivity <- transitivity(g, type = 'global', isolates = 'zero')
	dtout <- data.table(nEdges, nVertices, diameter, distance, density, transitivity)
	return(dtout)
}

getAssortativity <- function(dt){
	vertex_attributes <- melt(dt, id.var = names(dt)[3], value.name = 'vertex')
	vertex_attributes <- vertex_attributes[, .SD[1], 
		.SDcols = c('vertex', names(dt)[3]), by = vertex]
	g <- graph_from_data_frame(dt, directed = F, 
		vertices = vertex_attributes)
	assortativity <- assortativity_nominal(g, type = V(g)$from_print_order)
	return(assortativity)
}

images <- unique(cellNeighbours[['ImageNumber']])
getSubgraphs <- function(image, dt){
	image <- dt[ImageNumber == image]
	graphCols <- c('from', 'to')
	communityCols <- 'communities'
	image[,  eval(communityCols) := getCommunities(.SD), .SDcols = graphCols]
	properties <- c('nEdges', 'nVertices', 'diameter', 
		'distance', 'density', 'transitivity')
	image[, eval(properties) := getNetworkProperties(.SD), 
		.SDcols = graphCols, by = communities]
	image[, assortativity := getAssortativity(.SD), 
		.SDcols =c('from', 'to', 'from_print_order'), by = communities]
	image <- image[,.SD, .SDcols = c('ImageNumber', graphCols, 
		communityCols, properties, 'assortativity')]
	return(image)	
}

renameCommunityCols <- function(dt, type){
	communityCol <- 'communities'
	newname <- paste0(communityCol, '_', type)
	setnames(dt, communityCol, newname)
}

# compartment communities
mkGraphOut <- function(dt, type){
	images <- unique(dt[['ImageNumber']])		
	SubGraphs <- mclapply(images, getSubgraphs, dt, mc.cores = n_threads)
	out <- rbindlist(SubGraphs)
	renameCommunityCols(out, type)
	return(out)
}

# Split epithelial and non-epithelial cells per image, exclude perivascular cells.
tumour <- cellNeighbours[
	(from_is_epithelial == T & to_is_epithelial == T)
	& (from_is_vascular == F & to_is_vascular == F)
	]
tumour <- mkGraphOut(tumour, 'tumour')
stromal <- cellNeighbours[(from_is_epithelial == F & to_is_epithelial == F) 
	& (from_is_vascular == F & to_is_vascular == F)]
stromal <- mkGraphOut(stromal, 'stromal')

# two datasets out
# 1. Communities ids
toMerge <- function(dt){
	communityCol <- grep('communities', names(dt), value = T)
	toKeep <- c(communityCol, 'ImageNumber', 'from', 'to')
	dt <- dt[, .SD, .SDcols = toKeep]
	return(dt)
}
dts <- list(tumour = tumour, stroma = stromal)
toMerge <- lapply(dts, toMerge)
mergeTogether <- function(x, y){
	out <- merge(x, y, by = c('ImageNumber', 'from', 'to'), all = T)
}
allOut <- Reduce(mergeTogether, toMerge)

# 2. Network properties as list
toNetworkLevel <- function(dt){
	communityCol <- grep('communities', names(dt), value = T)
	idCols <- c('ImageNumber', communityCol)
	dt <- dt[, .SD[1], by = idCols]
	colstoKeep <- setdiff(names(dt), c('from', 'to'))
	dt <- dt[, .SD, .SDcols = colstoKeep]
	return(dt)
}
NetworkLevelDts <- mclapply(dts, toNetworkLevel, mc.cores = n_threads)

communities <- allOut
NetworkProperties <- NetworkLevelDts
rm(list = setdiff(ls(), c('cells', 'cellNeighbours', 'communities', 'NetworkProperties')))
communitiesAll <- communities

```

## Connectivity profiles
3.	Define connectivity profiles for all communities as vertex degree per phenotype


## Cluster communities
4.	Cluster with Ward’s method for hierarchical clustering


```{r}
#There are no train/validation labels available so we consider all cells for now. 
cells <- cells %>%
  add_column(isDiscovery = TRUE)

cellNeighbours <- cellNeighbours[, .SD, 
	.SDcols = c('ImageNumber', 'from', 'to', 'from_meta_description', 'to_meta_description')]

communities <- melt(communities, id.vars = c('ImageNumber', 'from', 'to'), 
	variable = 'type', value = 'communityID')
communities <- communities[!is.na(communityID)]
communities <- communities[grep('stromal', type)]

trainValID <- cells[, .(ImageNumber, isDiscovery, metabric_id)]
trainValID <- trainValID[, .SD[1], by = .(ImageNumber)]
communities <- merge(x = communities, y = trainValID, by = 'ImageNumber') 
setkeyv(communities, c('ImageNumber', 'communityID')) #mk unique community id
communities[, communityUID := as.integer(seq_len(.N) == 1L), by = .(ImageNumber, communityID)]
communities[, communityUID := cumsum(communityUID)]
communities <- merge(x = communities, y = cellNeighbours, by = c('ImageNumber', 'from', 'to'))
communities[, TotalInteractions := .N, by = communityUID]
communities[, nInteractions := asinh(.N), by = .(communityUID, to_meta_description)]
toCluster <- dcast(communities[, .SD[1], by = .(communityUID, to_meta_description)], 
	ImageNumber + communityID + communityUID + metabric_id + isDiscovery + TotalInteractions ~ to_meta_description, 
	value.var = 'nInteractions', fill = 0L)

```



```{r}
# Cell count threshold
minCount <- 20

mkCommunityModules <- function(Communities){
  #NOTE:Uncertain which variables this grep() function is trying to grasp --> Ep: epithelial, SI: stromal_immune
	# measureVars <- grep('Ep|SI', names(Communities), value = T)
	measureVars <- c("TotalInteractions")
  
	k <- 10 # n clusters	

	cat('Minimum interactions is:', minCount, '\n')
	cat('No of clusters is:', k, '\n')
	
	Communities <- Communities[TotalInteractions > minCount]
	Communities[, id := seq_len(.N)]

	getHCGroups <- function(dt, k) {
		dmat <- dist(dt, method = 'canberra')
		hc <- hclust(dmat, method = 'ward.D2')
		order <- hc$order
		groups <- stats::cutree(hc, k = k)
		return(list(groups = groups, order = order))
	}
	
	toHC <- as.matrix(Communities[, .SD, .SDcols = c(measureVars, 'id')], rownames = 'id')
	hcOut <- getHCGroups(toHC, k = k)
	Groups <- data.table(CommunityModuleID = hcOut$groups, 
		id = as.integer(names(hcOut$groups)))		
	Communities <- merge(Communities, Groups, by = 'id')
	Communities[, id := NULL]
	return(Communities)
}

TMEStructuresDiscov <- mkCommunityModules(toCluster[(isDiscovery)])
setnames(TMEStructuresDiscov, names(TMEStructuresDiscov), gsub(' ', '', names(TMEStructuresDiscov)))
setnames(TMEStructuresDiscov, 'CommunityModuleID', 'TMEStructure')
```
```{r}
ptLevel <- TMEStructuresDiscov[, .(ImageNumber, communityID, communityUID, TMEStructure)]
communitiesAll <- communitiesAll[, .(ImageNumber, from, to, communities_stromal)][!is.na(communities_stromal)]
setnames(communitiesAll, 'communities_stromal', 'communityID')
ptLevel <- merge(x = communitiesAll, y = ptLevel, by = c('ImageNumber', 'communityID'))
IDs <- cells[, .(ImageNumber, metabric_id)]
IDs <- IDs[, .SD[1], by = .(ImageNumber)]
ptLevel <- merge(x = ptLevel, y = IDs, by = 'ImageNumber') 
ptLevel[, TotalInteractions := .N, by = metabric_id]
ptLevel[,  nInteractionsPerStructure := .N, by = .(metabric_id, TMEStructure)]
ptLevel <- 
	ptLevel[, .SD[1], by = .(metabric_id, TMEStructure)][, 
	.SD, .SDcols = grep('from|to', names(ptLevel), invert = TRUE, value = TRUE)]
allCombinations <- as.data.table(expand.grid(ptLevel[, unique(metabric_id)], ptLevel[, unique(TMEStructure)])) #zeros
setnames(allCombinations, c('Var1', 'Var2'), c('metabric_id', 'TMEStructure'))
ptLevel <- merge(x = allCombinations, y = ptLevel, by = c('metabric_id', 'TMEStructure'), all.x = TRUE) 
# ptLevel <- setorderv(ptLevel, c("metabric_id", "ImageNumber"))
# ptLevel <- ptLevel[order("metabric_id", "ImageNumber", decreasing = FALSE)]
ptLevel[, ImageNumber := ImageNumber[1], by = metabric_id]
ptLevel[order(TotalInteractions, decreasing = FALSE), TotalInteractions := TotalInteractions[1], by = metabric_id]
ptLevel[is.na(nInteractionsPerStructure), nInteractionsPerStructure := 0L]


```


## Validation
5.	Validate reproducibility by training a random forest classifier on connectivity profiles

We skip this for now, because the train/validation labels are not known.


## Additional properties
6.	Compute number of vertices, number of edges, diameter, density, transitivity and assortativity and Shannon diversity by cell phenotype.

These are computed earlier already

## Visualization: figure 4b

```{r}
library(ComplexHeatmap)
library(latex2exp)
source(here("UtilityFunctions.R"))

cellTypes <- get_phenotypes()
dat <- copy(TMEStructuresDiscov)
cellVars = names(dat)[7:22]
dat[, eval(cellVars) := lapply(.SD, scale_clip), .SDcols = cellVars]
toHMatrix <- as.matrix(dat[, .SD, .SDcols = cellVars])
rownames(toHMatrix) <- dat[, communityUID]

celltypes_SI <- copy(cellTypes$stromal_immune)
names(celltypes_SI$description) <- gsub(" ", "", names(celltypes_SI$description))

mapColours <- data.frame(cellTypes$stromal_immune$description, gsub(" ", "", names(celltypes_SI$description)), names(cellTypes$stromal_immune$colours))
names(mapColours) <- c("id", "type", "colour")
mapColours = mapColours[match(colnames(toHMatrix),mapColours$type),]

reorderSubset <- function(subset){
	subset <- dat[TMEStructure == subset]
	HC <- getHC(as.matrix(subset[, .SD, .SDcols = cellVars]), 
		distance = 'euclidean', clustering = 'ward.D2')
	subset <- setNames(subset[, TMEStructure], subset[, communityUID])
	return(subset[HC$order])
}

reorderedSubsets <- mclapply(sortunique(dat[, TMEStructure]), reorderSubset, mc.cores = detectCores())
toHMatrix <- toHMatrix[names(unlist(reorderedSubsets)),]
toHMatrix <- t(toHMatrix)

CellCols <- HeatmapAnnotation(
	bar = rownames(toHMatrix),
	col = list(bar = setNames(mapColours$colour, mapColours$type)), 
	show_legend = F,
	show_annotation_name = F,
	which = 'row'
)

structures <- getStructures()
setkey(structures$TME$dat, 'Structure')
strColours <- setNames(structures$TME$dat[, colours], structures$TME$dat[, Structure])	
columnAnno <- HeatmapAnnotation(str = unlist(reorderedSubsets), 
	col = list(str = strColours), 
	which = 'column', show_annotation_name = FALSE,
	show_legend = FALSE
)
columnLabels <- setNames(structures$TME$dat[, Label], structures$TME$dat[, Structure])

col_levs <- 19
sval <- -2; eval <- 2
cols <- viridisLite::magma(col_levs)
distance <- abs((eval - sval)) / (col_levs - 1)
range <- seq(sval, eval, distance)
hmcols <- circlize::colorRamp2(range, cols)
legend_ti <- 'z score'
thin_line <- gpar(lwd = 0.5)
labsi <- gpar(fontsize = 15)
legend_ti <- "z-score"
breaks <- c(sval, 0, eval)

cm <- ColorMapping(name = legend_ti,
    col_fun = hmcols)

hmlegend <- list(
	object = cm, 
	# at = breaks, 
	legend_height = unit(20, 'mm'),
	grid_width = unit(3, 'mm'),
	title_gp = labsi,
	labels_gp = labsi,
	legend_gp = gpar(lwd = 0.1),
	border = rgb(0,0,0, alpha = 0)
)

rowLabels <- sapply(rownames(toHMatrix), TeX)
rowLabels[16] <- expression(paste('M', Phi, ' & granulocytes'))
fontSize <- 22

hm <- Heatmap(toHMatrix,
		col = hmcols,
		column_split = unlist(reorderedSubsets),
		column_gap = unit(2.5, 'mm'),
		row_labels = unlist(rowLabels),
		column_title = sapply(columnLabels, latex2exp::TeX),
		column_title_side = "bottom",
		row_title_gp = gpar(fontsize = fontSize),
		column_title_gp = gpar(fontsize = fontSize),
		column_title_rot = -90,
		row_names_side = 'left',
		left_annotation = CellCols,
		row_dend_side = 'right',
		column_dend_side = 'bottom',
		show_row_names = T,
		show_column_names = F,
		cluster_columns = F,
		cluster_rows = T,
		show_row_dend = F,
		show_column_dend = F,
		top_annotation = columnAnno,
		bottom_annotation = columnAnno,
		row_names_gp = gpar(fontsize = fontSize),
		row_dend_gp = thin_line,
		column_dend_gp = thin_line,
		heatmap_legend_param = hmlegend,
		raster_device = 'png',
		raster_quality = 25,
		heatmap_height = unit(20, 'cm')
	)

pdf(here("output/heatmap4c.pdf"), width = 20, height = 12)
draw(hm)
dev.off()


```


