---
title: "Danenberg Reproduction"
author: "Niek Brouwer"
date: "`r Sys.Date()`"
output: html_document
---

```{r introduction, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we aim to reproduce the results of the paper Breast tumor microenvironment structures are associated with genomic features and clinical outcome by Danenberg et al. (https://www.nature.com/articles/s41588-022-01041-y).
We focus on the description of the spatial organization of the TME as presented in figures 4b,d,f and its prognostic impact as seen in figure 6.

The separation and phenotype classification of single cells is not reproduced and this is retrieved from the original paper.

```{r libraries and settings}
#With R version 4.0
library(tidyverse)
library(ggplot2)
library(data.table)
library(parallel)
library(igraph)
library(here)
library(fst)
library(assertthat)
library(RColorBrewer)
library(latex2exp)
library(cowplot)

source(here("UtilityFunctions.R"))

# set seed
projectSeed <- 89230689
set.seed(projectSeed)

cells <- getCells()
cellNeighbours <- getCellNeighbours()
phenotypes <- get_phenotypes()
structures <- getStructures()
images <- unique(cellNeighbours[['ImageNumber']])
n_threads <- parallel::detectCores()
```

## Data discovery
To visualize the input datasets, we reproduce extended figure 2, showing cell type counts and proportions.

```{r data discovery}
cell_dat <- cells[, .(ImageNumber, ObjectNumber, is_epithelial, meta_description,
		print_order, meta_id, colours)]

# Get rectangularised proportion (or count) data
# Hierarchical cluster per image, order by count data
count_dat <- cell_dat[, ncells := .N, by = .(ImageNumber, meta_description)] 
count_dat <- dcast(count_dat, ImageNumber ~ meta_description, value.var = 'ncells')
hc_order_count <- getHC(count_dat[, -c('ImageNumber'), with = F])	
hc_order_count <- count_dat[hc_order_count$order, hc_order_count := seq_len(.N)][, 
	.(ImageNumber, hc_order_count)]
cell_dat <- merge(cell_dat, hc_order_count, by = 'ImageNumber')

# Hierarchical cluster order by proportion data
prop_dat <- copy(cell_dat[, ncells := .N, by = .(ImageNumber, meta_description)])
prop_dat[, totalcells := .N, by = .(ImageNumber, is_epithelial)] 
prop_dat[, proportion := (ncells / totalcells)]
prop_dat <- dcast(prop_dat, ImageNumber ~ meta_description, 
	value.var = 'proportion')
prop_dat[is.na(prop_dat)] <- 0
hc_order_prop <- getHC(prop_dat[, -c('ImageNumber'), with = F])
hc_order_prop <- prop_dat[hc_order_prop$order, 
	hc_order_prop := seq_len(.N)][, .(ImageNumber, hc_order_prop)]
cell_dat <- merge(cell_dat, hc_order_prop, by = 'ImageNumber')	

cell_dat[, tokeep := (seq_len(.N) == 1), by = .(ImageNumber, meta_description)]
cell_dat <- cell_dat[tokeep == T][, tokeep := NULL]
cell_dat[is_epithelial == T, facet_by := 'Epithelial'][is_epithelial == F,
	facet_by := 'TME']

# Legend and stack order
legend_colours <- c(names(phenotypes$epithelial$colours),
	names(phenotypes$stromal_immune$colours))
legend_labels <- c(names(phenotypes$epithelial$description),
	names(phenotypes$stromal_immune$description))
legend_labels <- lapply(legend_labels, TeX)

cell_dat$colours <-
	factor(cell_dat$colours, levels = legend_colours, 
		ordered = T)

# Styling function
mkplot <- function(proportion = F){
	
	if (proportion){
		stack <- 'fill'
		xvar <- 'hc_order_prop'
		ylabs <- c('0', '0.5', '1')
		ybreaks <- c(0, 0.5, 1)
		yti <- 'Proportion'
	} else{
		stack <- 'stack'
		xvar <- 'hc_order_count'
		ylabs <- c('0', '4k', '8k')
		ybreaks <- c(0, 4e3, 8e3)
		yti <- 'Number of cells'
	}

	xmax <- cell_dat[, max(get(xvar))]

	p <- 
		ggplot() +
		geom_bar(data = cell_dat,
			aes(x = get(xvar),
				y = ncells,
				fill = colours),
			position = stack,
			stat = 'identity', width = 1
			) +
		scale_fill_identity(
			'Cell phenotype',
			breaks = legend_colours,
			labels = legend_labels,
			guide = 'legend'
			) +
		facet_wrap(~facet_by, nrow = 2) +
		scale_x_continuous(
			expand = c(0,0), 
			breaks = c(1, xmax)) +
		scale_y_continuous(
			expand = c(0,0),
			labels = ylabs,
			breaks = ybreaks
			) +
		theme_prj() +
		theme(
			plot.margin = unit(rep(0.5, 4), 'cm'),
			axis.line = element_blank(),
			axis.ticks = element_blank(),
			axis.text = element_text(size = 9),
			axis.title = element_text(size = 9),
			axis.title.y = element_text(angle = 90),
			strip.text  = element_text(size = 9, margin = margin(0,0,0,0, unit = 'pt')),
			strip.background = element_rect(fill = 'lightgrey', colour = 'lightgrey'),
			legend.position = 'bottom',
			legend.text = element_text(size = 9),
			legend.key.size = unit(0.5, 'line'),
			legend.title = element_text(size = 11)
			) +
		guides(fill = guide_legend(
			title.position = 'top')) +
		labs(x = 'Tissue image', y = yti)

}	

counts <- mkplot(proportion = F)
proportions <- mkplot(proportion = T)

counts <- extract_gglegend(counts)
legend <- counts$legend
counts <- counts$plot
proportions <- extract_gglegend(proportions)
proportions <- proportions$plot
p <- plot_grid(counts, proportions, ncol = 1, align = 'v')
p <- plot_grid(p, legend, ncol = 1, rel_heights = c(6, 1))
outfile <- file.path(here('output/stacked_bar_CountsProportionsCombined.pdf'))
ggsave(outfile, p, height = 7, width = 8)

```



## Community detection
The TME is represented as a graph to find densely connected cell clusters using a community-detection algorithm. number of edges, number of vertices, diameter, density, transitvity and assortativity is calculated for the resulting communities.

```{r community detection and property calculation}

# Communiy-detection algorithm using random walks
getCommunities <- function(dt){
	g <- graph_from_data_frame(dt, directed = F)
	subgraphs <- cluster_walktrap(g)
	assignments <- membership(subgraphs)
	matchOrder <- match(dt[[1]], as.numeric(names(assignments)))	
	communities <- assignments[matchOrder]
	return(data.table(communities))
}

# Retrieve network properties of individual communities for later analyses
getNetworkProperties <- function(dt){
	g <- graph_from_data_frame(dt, directed = F)
	# n of edges
	nEdges <- ecount(g)
	# n of vertices
	nVertices <- vcount(g)
	# length of the longest path (in number of edges) between two nodes
	diameter <- diameter(g) 
	# average number of edges between any two nodes in the network.
	distance <- mean_distance(g)
	# ratio of the number of edges and the number of possible edges
	density <- edge_density(g)
	# probability that the adjacent vertices of a vertex are connected. 
	transitivity <- transitivity(g, type = 'global', isolates = 'zero')
	dtout <- data.table(nEdges, nVertices, diameter, distance, density, transitivity)
	return(dtout)
}

# Compute assortativity of subgraphs
getAssortativity <- function(dt){
	vertex_attributes <- melt(dt, id.var = names(dt)[3], value.name = 'vertex')
	vertex_attributes <- vertex_attributes[, .SD[1], 
		.SDcols = c('vertex', names(dt)[3]), by = vertex]
	g <- graph_from_data_frame(dt, directed = F, 
		vertices = vertex_attributes)
	assortativity <- assortativity_nominal(g, type = V(g)$from_print_order)
	return(assortativity)
}

getSubgraphs <- function(image, dt){
	image <- dt[ImageNumber == image]
	graphCols <- c('from', 'to')
	# Find communities
	communityCols <- 'communities'
	image[,  eval(communityCols) := getCommunities(.SD), .SDcols = graphCols]
	# Compute properties
	properties <- c('nEdges', 'nVertices', 'diameter', 
		'distance', 'density', 'transitivity')
	image[, eval(properties) := getNetworkProperties(.SD), 
		.SDcols = graphCols, by = communities]
	image[, assortativity := getAssortativity(.SD), 
		.SDcols =c('from', 'to', 'from_print_order'), by = communities]
	image <- image[,.SD, .SDcols = c('ImageNumber', graphCols, 
		communityCols, properties, 'assortativity')]
	return(image)	
}

renameCommunityCols <- function(dt, type){
	communityCol <- 'communities'
	newname <- paste0(communityCol, '_', type)
	setnames(dt, communityCol, newname)
}

# compartment communities
mkGraphOut <- function(dt, type){
	images <- unique(dt[['ImageNumber']])		
	SubGraphs <- mclapply(images, getSubgraphs, dt, mc.cores = n_threads)
	out <- rbindlist(SubGraphs)
	renameCommunityCols(out, type)
	return(out)
}

# Split epithelial and non-epithelial cells per image, exclude perivascular cells.
tumour <- cellNeighbours[
	(from_is_epithelial == T & to_is_epithelial == T)
	& (from_is_vascular == F & to_is_vascular == F)
	]
tumour <- mkGraphOut(tumour, 'tumour')
stromal <- cellNeighbours[(from_is_epithelial == F & to_is_epithelial == F) 
	& (from_is_vascular == F & to_is_vascular == F)]
stromal <- mkGraphOut(stromal, 'stromal')

# Merge two results
# 1. Communities ids
toMerge <- function(dt){
	communityCol <- grep('communities', names(dt), value = T)
	toKeep <- c(communityCol, 'ImageNumber', 'from', 'to')
	dt <- dt[, .SD, .SDcols = toKeep]
	return(dt)
}
dts <- list(tumour = tumour, stroma = stromal)
toMerge <- lapply(dts, toMerge)
mergeTogether <- function(x, y){
	out <- merge(x, y, by = c('ImageNumber', 'from', 'to'), all = T)
}
allOut <- Reduce(mergeTogether, toMerge)

# 2. Network properties as list
toNetworkLevel <- function(dt){
	communityCol <- grep('communities', names(dt), value = T)
	idCols <- c('ImageNumber', communityCol)
	dt <- dt[, .SD[1], by = idCols]
	colstoKeep <- setdiff(names(dt), c('from', 'to'))
	dt <- dt[, .SD, .SDcols = colstoKeep]
	return(dt)
}
NetworkLevelDts <- mclapply(dts, toNetworkLevel, mc.cores = n_threads)

# Collect result and clean intermediate results
communities <- copy(allOut)
NetworkProperties <- copy(NetworkLevelDts)
rm(list = setdiff(ls(), c('communities', 'NetworkProperties')))
```


## Cluster communities
connectivity profiles (phenotype composition and vertex degree) are computed for the resulting communities.

```{r setup clustering}
# Retrieve input datasets
source(here("UtilityFunctions.R"))
cells <- getCells()
cellNeighbours <- getCellNeighbours()
phenotypes <- get_phenotypes()
communitiesResult <- copy(communities) #save a version of the original communities

cellNeighbours <- cellNeighbours[, .SD, 
	.SDcols = c('ImageNumber', 'from', 'to', 'from_meta_id', 'to_meta_id')]

communities <- melt(communities, id.vars = c('ImageNumber', 'from', 'to'), 
	variable = 'type', value = 'communityID')
communities <- communities[!is.na(communityID)]
communities <- communities[grep('stromal', type)]

trainValID <- getCells()[, .(ImageNumber, isDiscovery, metabric_id)]
trainValID <- trainValID[, .SD[1], by = .(ImageNumber)]
communities <- merge(x = communities, y = trainValID, by = 'ImageNumber') 
setkeyv(communities, c('ImageNumber', 'communityID')) #mk unique community id
communities[, communityUID := as.integer(seq_len(.N) == 1L), by = .(ImageNumber, communityID)]
communities[, communityUID := cumsum(communityUID)]
communities <- merge(x = communities, y = cellNeighbours, by = c('ImageNumber', 'from', 'to'))
# Compute total interactions of a community
communities[, TotalInteractions := .N, by = communityUID]
# Compute interactions per cell type
communities[, nInteractions := asinh(.N), by = .(communityUID, to_meta_id)]
toCluster <- dcast(communities[, .SD[1], by = .(communityUID, to_meta_id)], 
	ImageNumber + communityID + communityUID + metabric_id + isDiscovery + TotalInteractions ~ to_meta_id, 
	value.var = 'nInteractions', fill = 0L)
```

The communities are clustered with Ward’s method for hierarchical clustering.

```{r clustering}
# Cell count threshold
minCount <- 20

mkCommunityModules <- function(Communities){
	measureVars <- grep('Ep|SI', names(Communities), value = T)
	
	k <- 10 # n clusters	

	cat('Minimum interactions is:', minCount, '\n')
	cat('No of clusters is:', k, '\n')
	
	Communities <- Communities[TotalInteractions > minCount]
	Communities[, id := seq_len(.N)]

	# Hierarchical clustering
	getHCGroups <- function(dt, k) {
		dmat <- dist(dt, method = 'canberra')
		hc <- hclust(dmat, method = 'ward.D2')
		order <- hc$order
		groups <- stats::cutree(hc, k = k)
		return(list(groups = groups, order = order))
	}
	
	toHC <- as.matrix(Communities[, .SD, .SDcols = c(measureVars, 'id')], rownames = 'id')
	hcOut <- getHCGroups(toHC, k = k)
	Groups <- data.table(CommunityModuleID = hcOut$groups, 
		id = as.integer(names(hcOut$groups)))		
	Communities <- merge(Communities, Groups, by = 'id')
	Communities[, id := NULL]
	return(Communities)
}

TMEStructuresDiscov <- mkCommunityModules(toCluster[(isDiscovery)])
setnames(TMEStructuresDiscov, names(TMEStructuresDiscov), gsub(' ', '', names(TMEStructuresDiscov)))
setnames(TMEStructuresDiscov, 'CommunityModuleID', 'TMEStructure')
```

## Validation
Reproducibility is validated by training a random forest classifier on connectivity profiles.

```{r validation}
library(diffdf)

RFCCols <- c('TMEStructure', grep('SI', names(TMEStructuresDiscov), value = TRUE))
#Train random forest model on cell type proportions
RFCModel <- randomForest::randomForest(formula = as.factor(TMEStructure) ~ ., 
	data = TMEStructuresDiscov[, .SD, .SDcols = RFCCols],
	importance = TRUE, mtry = 12)
RFCModel

#Select validation set
toClusterValidation <- toCluster[!(isDiscovery)][TotalInteractions > minCount]
oldNames <- grep('SI', names(toClusterValidation), value = TRUE)
newNames <- gsub(' ', '', oldNames)
setnames(toClusterValidation, oldNames, newNames)
# Predict community label
predictions <- predict(RFCModel, 
	toClusterValidation[, .SD, 
	.SDcols = grep('SI', names(toClusterValidation), value = TRUE)])
toClusterValidation[, TMEStructure := predictions] 

#Combine datasets
TMEStructuresDiscov[, isDiscovery := TRUE]
toClusterValidation[, isDiscovery := FALSE]
TMEStructures <- rbindlist(list(TMEStructuresDiscov, toClusterValidation))
outfile <- here('scratch/TMEStructures.fst')
write_fst(TMEStructures, outfile, compress = 100) 

cellNeighbours <- getCellNeighbours()
checkIDs <- cellNeighbours[, .(ImageNumber, communityID, TMEStructure)]
checkIDs <- checkIDs[, .SD[1], by = .(ImageNumber, communityID)]
checkIDs <- na.omit(checkIDs)
setkeyv(checkIDs, names(checkIDs))
setkeyv(TMEStructures, names(checkIDs))

# Print wrongly predicted communities
issues <- diffdf(TMEStructures[, .SD, .SDcols = names(checkIDs)], checkIDs)
false_discovery <- TMEStructures[issues$VarDiff_TMEStructure$..ROWNUMBER..,]
false_discovery$predict_id <- as.character(issues$VarDiff_TMEStructure$COMPARE)
false_discovery$TMEStructure <- as.character(false_discovery$TMEStructure)
false_discovery <- false_discovery %>% rename("true_id" = "TMEStructure")

# false_discovery[,23:24]
# 
# structures <- getStructures()
# 
# cell_labels <- data.frame(as.character(structures$TME$dat$Structure), as.character(structures$TME$dat$Label))
# setnames(cell_labels, c("id", "structure"))
# 
# 
# false_discovery <- left_join(false_discovery, cell_labels, by=c("true_id" = "id"))
# false_discovery <- false_discovery %>% rename("true_structure" = "structure")
# false_discovery <- left_join(false_discovery, cell_labels, by=c("predict_id" = "id"))
# false_discovery <- false_discovery %>% rename("predict_structure" = "structure")
# 
# library("caret")
# confusionMatrix(as.factor(false_discovery$true_structure), as.factor(false_discovery$predict_structure))
```
The resulting communities are aggregated per patient

```{r cluster aggregation}

ptLevel <- TMEStructures[, .(ImageNumber, communityID, communityUID, TMEStructure)]
#Select TME communities
communitiesAll <- communitiesResult[, .(ImageNumber, from, to, communities_stromal)][!is.na(communities_stromal)]
setnames(communitiesAll, 'communities_stromal', 'communityID')
ptLevel <- merge(x = communitiesAll, y = ptLevel, by = c('ImageNumber', 'communityID'))
IDs <- getCells()[, .(ImageNumber, metabric_id)]
IDs <- IDs[, .SD[1], by = .(ImageNumber)]
ptLevel <- merge(x = ptLevel, y = IDs, by = 'ImageNumber')
#Compute interactions (vertex degree)
ptLevel[, TotalInteractions := .N, by = metabric_id]
ptLevel[,  nInteractionsPerStructure := .N, by = .(metabric_id, TMEStructure)]
ptLevel <- 
	ptLevel[, .SD[1], by = .(metabric_id, TMEStructure)][, 
	.SD, .SDcols = grep('from|to', names(ptLevel), invert = TRUE, value = TRUE)]
allCombinations <- adt(expand.grid(ptLevel[, unique(metabric_id)], ptLevel[, unique(TMEStructure)])) #zeros
setnames(allCombinations, c('Var1', 'Var2'), c('metabric_id', 'TMEStructure'))
ptLevel <- merge(x = allCombinations, y = ptLevel, by = c('metabric_id', 'TMEStructure'), all.x = TRUE) 
ptLevel <- ptLevel[order(metabric_id, ImageNumber, decreasing = FALSE)]
ptLevel[, ImageNumber := ImageNumber[1], by = metabric_id]
ptLevel[order(TotalInteractions, decreasing = FALSE), TotalInteractions := TotalInteractions[1], by = metabric_id]
ptLevel[is.na(nInteractionsPerStructure), nInteractionsPerStructure := 0L]
outfile <- here('scratch/ptLeveLTMEInteractions.fst')
write_fst(ptLevel, outfile, compress = 100)

#Clean intermediate results
rm(list = setdiff(ls(), c('communitiesResult', 'NetworkProperties', 'TMEStructures', 'TMEStructuresDiscov')))

write_rds(NetworkProperties, here('scratch/NetworkProperties.rds'))


```


## Visualization: figure 4b Heatmap



```{r heatmap 4b}
library(ComplexHeatmap)
library(latex2exp)
source(here("UtilityFunctions.R"))

toHM <- function(dat, fileToSave, height = NULL, isDiscovery) {
	cellTypes <- get_phenotypes()
	dat <- copy(dat)
	cellVars <- grep('SI', names(dat), value = TRUE)
	dat[, eval(cellVars) := lapply(.SD, scale_clip), .SDcols = cellVars]
	toHMatrix <- as.matrix(dat[, .SD, .SDcols = cellVars])
	rownames(toHMatrix) <- dat[, communityUID]
	mapColNamesIndex <- as.integer(gsub('SI', '', colnames(toHMatrix)))
	mapColNames <- names(cellTypes$stromal_immune$description[mapColNamesIndex])
	colnames(toHMatrix) <- mapColNames
	mapColours <- cellTypes$stromal_immune$colours[mapColNamesIndex]
	mapColours <- setNames(names(mapColours), mapColNames)

	reorderSubset <- function(subset){
		subset <- dat[TMEStructure == subset]
		HC <- getHC(as.matrix(subset[, .SD, .SDcols = grep('SI', names(subset), value = TRUE)]), 
			distance = 'euclidean', clustering = 'ward.D2')
		subset <- setNames(subset[, TMEStructure], subset[, communityUID])
		return(subset[HC$order])
	}
	reorderedSubsets <- mclapply(sortunique(dat[, TMEStructure]), reorderSubset, mc.cores = detectCores())
	toHMatrix <- toHMatrix[names(unlist(reorderedSubsets)),]
	toHMatrix <- t(toHMatrix)

	cat('Is this discovery data?', isDiscovery, '\n')
	cat('There are', ncol(toHMatrix), 'columns\n')

	CellCols <- HeatmapAnnotation(
		bar = rownames(toHMatrix),
		col = list(bar = mapColours), 
		show_legend = F,
		show_annotation_name = F,
		which = 'row'
	)

	structures <- getStructures()
	setkey(structures$TME$dat, 'Structure')
	strColours <- setNames(structures$TME$dat[, colours], structures$TME$dat[, Structure])	
	columnAnno <- HeatmapAnnotation(str = unlist(reorderedSubsets), 
		col = list(str = strColours), 
		which = 'column', show_annotation_name = FALSE,
		show_legend = FALSE
	)
	columnLabels <- setNames(structures$TME$dat[, Label], structures$TME$dat[, Structure])	

	col_levs <- 19
	sval <- -2; eval <- 2
	cols <- viridisLite::magma(col_levs)
	distance <- abs((eval - sval)) / (col_levs - 1)
	range <- seq(sval, eval, distance)
	hmcols <- circlize::colorRamp2(range, cols)
	legend_ti <- 'z score'
	thin_line <- gpar(lwd = 0.5)
	labsi <- gpar(fontsize = 15)
	legend_ti <- "z-score"
	breaks <- c(sval, 0, eval)

	cm <- ColorMapping(name = legend_ti,
	    col_fun = hmcols)

	hmlegend <- list(
		object = cm, 
		# at = breaks, 
		legend_height = unit(20, 'mm'),
		grid_width = unit(3, 'mm'),
		title_gp = labsi,
		labels_gp = labsi,
		legend_gp = gpar(lwd = 0.1),
		border = rgb(0,0,0, alpha = 0)
	)

	rowLabels <- sapply(rownames(toHMatrix), TeX)
	# rowLabels[16] <- expression(paste('M', Phi, ' & granulocytes'))
	fontSize <- 22

	hm <- Heatmap(toHMatrix,
		col = hmcols,
		column_split = unlist(reorderedSubsets),
		column_gap = unit(2.5, 'mm'),
		row_labels = unlist(rowLabels),
		column_title = sapply(columnLabels, latex2exp::TeX),
		column_title_side = "bottom",
		row_title_gp = gpar(fontsize = fontSize),
		column_title_gp = gpar(fontsize = fontSize),
		column_title_rot = 90,
		row_names_side = 'left',
		left_annotation = CellCols,
		row_dend_side = 'right',
		column_dend_side = 'bottom',
		show_row_names = T,
		show_column_names = F,
		cluster_columns = F,
		cluster_rows = T,
		show_row_dend = F,
		show_column_dend = F,
		top_annotation = columnAnno,
		bottom_annotation = columnAnno,
		row_names_gp = gpar(fontsize = fontSize),
		row_dend_gp = thin_line,
		column_dend_gp = thin_line,
		heatmap_legend_param = hmlegend,
		raster_device = 'png',
		raster_quality = 25,
		heatmap_height = unit(20, 'cm')
	)

	if(is.null(height)){
		w <- 20; h <- 12
		if(is.null(fileToSave)) pdf('test.pdf', width = w, height = h)
		else pdf(fileToSave, width = w, height = h)	
		draw(hm)
		dev.off()
	}
	return(hm)

}

DiscovHM <- toHM(TMEStructures[(isDiscovery)], isDiscovery = TRUE, 
	fileToSave = file.path(here('output/TMEStrDiscoveryHM.pdf')))
# 
ValHM <- toHM(TMEStructures[!(isDiscovery)], isDiscovery = FALSE,
	fileToSave = file.path(here('output/TMEStrValidationHM.pdf')))

combined <- DiscovHM + ValHM
pdf(file.path(here('output/combinedTMEStrHM.pdf')), width = 20, height = 12)
draw(combined)
dev.off()

# Clean intermediate results
rm(list = setdiff(ls(), c('communitiesResult', 'NetworkProperties', 'TMEStructures', 'TMEStructuresDiscov')))

```

## Visualization: figure 4d counts and diversity

```{r plot 4d}
library(vegan)
source(here("UtilityFunctions.R"))

communitiesIDs <- communitiesResult[,c('ImageNumber', 'from', 'to', 'communities_stromal')]
communitiesIDs <- communitiesIDs[!is.na(communities_stromal)]
setnames(communitiesIDs, 'communities_stromal', 'communityID')
communityTMEMapping <- here('scratch/TMEStructures.fst')
colstokeep <- read_fst(communityTMEMapping, as.data.table = T, from = 1, to = 1)
communityTMEMapping <- read_fst(communityTMEMapping, as.data.table = T,
	columns = grep('SI', names(colstokeep), invert = TRUE, value = TRUE))
structures <- getStructures()
structures$TME$dat$Structure <- as.character(structures$TME$dat$Structure)
communityTMEMapping <- merge(x = communityTMEMapping, y = structures$TME$dat, 
	by.x = 'TMEStructure', by.y = 'Structure')

oldNames <- c('Label', 'colours', 'print_order')
newNames <- paste0('TMEStr', oldNames)
setnames(communityTMEMapping, oldNames, newNames)
communityIDs <- merge(x = communitiesIDs, y = communityTMEMapping, 
	by= c('ImageNumber', 'communityID')) 
setkeyv(communityIDs, c('ImageNumber', 'communityID', 'from'))

nwProperties <- copy(NetworkProperties)
nwProperties <- nwProperties[['stroma']]
setnames(nwProperties, 'communities_stromal', 'communityID')
communityIDs <- merge(x = communityIDs, y = nwProperties[, .(ImageNumber, communityID, nVertices)], 
	by = c('ImageNumber', 'communityID')) 

cellDat <- getCells()[(isTumour)]
communityIDs <- merge(x = communityIDs, y = cellDat, 
	by.x = c('ImageNumber', 'from'), by.y = c('ImageNumber', 'ObjectNumber')) 
communityIDs[, uniqueCell := (seq_len(.N) == 1L), by = .(ImageNumber, communityID, from)]
communityIDs[(uniqueCell), shannonDiversity := vegan::diversity(TMEStrprint_order), by = .(communityUID)]

toPlot <- communityIDs[, .SD[1], by = communityUID]
toPlot[, TotalCommunities := .N, by = TMEStructure]
toPlot[, medianNVertices := median(as.numeric(nVertices)), by = TMEStructure]
toPlot <- toPlot[order(medianNVertices, TMEStructure)]
toPlot[, yaxis := as.integer(seq_len(.N) == 1L), by = TMEStructure]
toPlot[, yaxis := cumsum(yaxis)]
toPlot <- toPlot[order(yaxis, decreasing = TRUE)]

toPlot[, yaxisRev := as.integer(seq_len(.N) == 1L), by = yaxis]
toPlot[, yaxisRev := cumsum(yaxisRev)]
toPlot[, yaxisRev := factor(yaxisRev, 
	levels = unique(yaxisRev), labels = unique(TMEStrLabel), ordered = TRUE)]
toPlot <- melt(toPlot, id.vars = c('yaxisRev', 'communityUID', 'TMEStrcolours', 'TotalCommunities'), 
	measure.vars = c('nVertices', 'shannonDiversity'))
toPlot[, variable := ifelse(grepl('nVertices', variable), 'Cell count', 'Shannon diversity')]

toPlot$yaxisRev <- factor(toPlot$yaxisRev,
    levels = c("Suppressed expansion", "TLS-like", "Vascular stroma", "$PDPN^{+}$ active stroma","Active IR",
               "$FSP1^{+}$ enriched$", "Granulocyte enriched", "APC enriched",  "$CD8^{+}$ & $macrophages$",
                                   "Active stroma"),ordered = TRUE)

lineW <- 0.2
box <- ggplot() +
	geom_boxplot(data = toPlot, 
		aes(x = yaxisRev, y = value, fill = TMEStrcolours),
		outlier.size = 0.1, size = lineW) +
	scale_fill_identity() +
	scale_x_discrete(labels = sapply(as.character(toPlot$yaxisRev), TeX)) +
	facet_grid(cols = vars(variable), 
		scales = 'free', space = 'fixed') + 
	theme_prj(base_line_size = lineW) +
	theme(axis.title.x = element_blank(),
		plot.margin = unit(c(0,0,0,0), 'npc')) +
	coord_flip() +
	labs(x = 'TME structure')

toPlot[variable == 'Cell count', countfacet := factor(1, levels = 1, 
	labels = expression('n communities'), ordered = TRUE)]

counts <- ggplot(data = toPlot[, .SD[1], by = yaxisRev]) +
	geom_col(aes(x = yaxisRev, y = TotalCommunities), fill = 'lightgrey') +
	geom_text(aes(x = yaxisRev, y = TotalCommunities, label = TotalCommunities),
		hjust = 0, nudge_y = 50) +
	theme_prj(base_line_size = lineW) +
	theme(axis.text.y = element_blank(),
		axis.text.x = element_text(colour = 'white'),
		axis.title = element_blank(),
		axis.line.y = element_blank(),
		axis.ticks.y = element_blank(),
		axis.line.x = element_line(colour = 'white'),
		axis.ticks.x = element_line(colour = 'white'),
		plot.margin = unit(c(0,0.4,0,0), 'npc')) +
	coord_flip(clip = 'off') +
	scale_x_discrete(labels = sapply(as.character(toPlot$yaxisRev), TeX)) +
	scale_y_continuous(expand = c(0,0), limits = c(0,4000)) +
	facet_grid(cols = vars(countfacet)) 


p <- plot_grid(box, counts, nrow = 1, rel_widths = c(1, 0.4))
save_plot(here('output/TMEStructuresVCellCountShannonBox.pdf'), p,
	base_height = 1.9, base_width = 5.4)

# Clean intermediate results
rm(list = setdiff(ls(), c('communitiesResult', 'NetworkProperties', 'TMEStructures', 'TMEStructuresDiscov')))
```



## Visualization: figure 4f vertex degree

```{r plot 4f}
library('igraph')
source(here("UtilityFunctions.R"))

cellNeighbours <- getCellNeighbours()
cellNeighbours <- cellNeighbours[, .SD, 
	.SDcols = setdiff(names(cellNeighbours), c('metabric_id', 'TMEStructure', 'communityID', 'isDiscovery'))]
communitiesIDs <- communitiesResult[,c('ImageNumber', 'from', 'to', 'communities_stromal')]
communitiesIDs <- communitiesIDs[!is.na(communities_stromal)]
setnames(communitiesIDs, 'communities_stromal', 'communityID')
communityTMEMapping <- here('scratch/TMEStructures.fst')
colstokeep <- read_fst(communityTMEMapping, as.data.table = T, from = 1, to = 1)
communityTMEMapping <- read_fst(communityTMEMapping, as.data.table = T,
	columns = grep('SI', names(colstokeep), invert = TRUE, value = TRUE))
communityIDs <- merge(x = communitiesIDs, y = communityTMEMapping, 
	by = c('ImageNumber', 'communityID')) 
cellNeighbours <- merge(x = cellNeighbours, y = communityIDs, 
	by = c('ImageNumber', 'from', 'to')) 

mkDegrees <- function(graph_dt){
	degrees <- graph_from_data_frame(graph_dt[, .(from, to)], directed = F)
	degrees <- simplify(degrees)
	nVertices <- length(V(degrees))
	degrees <- degree(degrees, loops = F)
	degrees_dt <- data.table(from = as.integer(names(degrees)), 
		degree = degrees, nVertices = nVertices)
	graph <- merge(x = graph_dt, y = degrees_dt, by = 'from') 
	return(graph)
}

Degrees <- split(cellNeighbours, by = 'communityUID')
Degrees <- mclapply(Degrees, mkDegrees, mc.cores = detectCores())
Degrees <- rbindlist(Degrees)
stopifnot(Degrees[is.na(degree), .N] == 0L)
maxDegreeClip <- 9
Degrees[, degree := ifelse(degree > maxDegreeClip, maxDegreeClip, degree)]
Degrees[, totalDegree := .N, by = .(TMEStructure, degree)]
Degrees[, proportion := .N / totalDegree, by = .(TMEStructure, degree, to_meta_description)]

toPlot <- Degrees[, .SD[1], by = .(TMEStructure, degree, to_meta_description)]
cellPhenotypes <- get_phenotypes()
filllabs <- names(cellPhenotypes$stromal_immune$description)
fillcols <- cellPhenotypes$stromal_immune$colours
toPlot[, fillColOrder := factor(to_print_order, 
	levels = sortunique(to_print_order), 
	labels = names(fillcols),  ordered = TRUE)]

structures <- getStructures()
toMapVar <- structures$TME$dat[order(print_order)]
mapVar <- setNames(toMapVar[, print_order], toMapVar[,Label])
toPlot[, TMEStructureNo := mapVar[TMEStructure]]
labels <- structures$TME$dat$Label
facetlabs <- unlist(lapply(labels, TeX), use.names = F)
printMap <- setNames(structures$TME$dat[,print_order], structures$TME$dat[,Structure])
toPlot[, facet_by := printMap[as.character(TMEStructure)]]
toPlot[, facet_by := factor(facet_by, 
	levels = as.integer(mapVar), labels = facetlabs, ordered = TRUE)]

xbreaks <- toPlot[, sortunique(degree)]
xlabs <- as.character(xbreaks)
xlabs <- gsub(as.character(maxDegreeClip), paste0(maxDegreeClip, '+'), xlabs)
legend_colours <- names(fillcols)
legend_labels <- unlist(lapply(filllabs, TeX), use.names = FALSE)

mkStackedArea <- function(dat){
	p <- ggplot() +
	geom_area(data = dat, 
		aes(x = degree, y = proportion, fill = fillColOrder), size = 1) +
	scale_y_continuous(expand = c(0,0), breaks = c(0,1), labels = c('0', '1')) +
	scale_x_continuous(expand = c(0,0), breaks = xbreaks, labels = xlabs) +
	scale_fill_identity(
		'Cell phenotype',
		breaks = legend_colours,
		labels = legend_labels,
		guide = 'legend'
	) +		
	theme_prj(base_line = 0.25) +
	facet_wrap(vars(facet_by), nrow = 2) +
	labs(x = 'Vertex degree (cell interactions)', y = 'Proportion') +
	guides(fill = guide_legend(title.position = 'top', ncol = 1, 
		keywidth = 0.5, keyheight = 0.5, label.position = 'left', label.hjust = 1)) +
	theme(
		axis.text = element_text(size = 5),
		strip.background = element_blank(), 
		strip.text = element_text(size = 6.25, margin = margin(t = 0, b = 0)),
		panel.spacing = unit(1, 'mm'),
		legend.position = 'left', legend.title.align = 1)
	return(p)
}

plotOut <- mkStackedArea(dat = toPlot)
save_plot(here('output/TMEStructuresVertexDegreeStackedArea.pdf'), plotOut, base_width = 7.35)
```

```{r}
# cellCentric # Ext Data 7
cellDegrees <- getCellNeighbours()
cellDegrees <- cellDegrees[from_is_epithelial == to_is_epithelial]
cellDegrees <- cellDegrees[!(from_is_epithelial)]
cellDegrees[, degree := .N, by = .(ImageNumber, from)]

maxDegreeClip <- 9
cellDegrees[, degree := ifelse(degree > maxDegreeClip, maxDegreeClip, degree)]

cellDegrees[, totalDegree := .N, by = .(from_meta_description, degree)]
cellDegrees[, proportion := .N / totalDegree, by = .(from_meta_description, degree, to_meta_description)]

toPlot <- cellDegrees[, .SD[1], by = .(from_meta_description, degree, to_meta_description)]
cellPhenotypes <- get_phenotypes()
filllabs <- names(cellPhenotypes$stromal_immune$description)
fillcols <- names(cellPhenotypes$stromal_immune$colours)
toPlot[, fillColOrder := factor(to_print_order,
	levels = sortunique(to_print_order),
	labels = fillcols,  ordered = TRUE)]

facetlabs <- names(cellPhenotypes$stromal_immune$description)
facetlabs <- unlist(lapply(facetlabs, TeX), use.names = FALSE)
# facetlabs[9] <- expression(paste('M', Phi, ' & granulocytes'))
toPlot[, facet_by := factor(from_print_order,
	levels = sortunique(from_print_order), labels = facetlabs, ordered = TRUE)]

xbreaks <- toPlot[, sortunique(degree)]
xlabs <- as.character(xbreaks)
xlabs <- gsub(as.character(maxDegreeClip), paste0(maxDegreeClip, '+'), xlabs)
legend_colours <- fillcols
legend_labels <- unlist(lapply(filllabs, TeX), use.names = FALSE)


plotOut <- mkStackedArea(dat = toPlot)
save_plot(here('output/TMECellCentricVertexDegreeStackedArea.pdf'), plotOut, base_width = 11)

#Clean intermediate results
# rm(list = setdiff(ls(), c('communitiesResult', 'NetworkProperties', 'TMEStructures', 'TMEStructuresDiscov')))
```


