---
title: "Data_exploration"
author: "Niek Brouwer"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: "hide"
params:
  save_files:
    value: FALSE
    choices:
      - FALSE
      - TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data exploration
In this document we explore the METABRIC dataset.

```{r, include=FALSE}
library(ComplexHeatmap)
library(tidyHeatmap)
library(here)
library(ggplot2)
library(tidyverse)
library(fst)
source(here("UtilityFunctions.R"))
source(here("MEP_UtilityFunctions.R"))
library(patchwork)
library(cowplot)
library(caret)
library(data.table)

cells <- getCells()
cell_counts <- getCellCounts(cells)
clinical_data <- getClinical()
cell_occurences_combinations <- getCombinationCounts(cells)
```

## Basic information

```{r}
print(paste('Number of samples: ', length(unique(cells$ImageNumber))))
print(paste('Number of cell types: ', length(unique(cells$meta_description))))
```

```{r}
ggplot(data= cells %>% count(ImageNumber)) + geom_bar(aes(y=n, x=ImageNumber), stat='identity') + theme_bw() + ggtitle('Total counts per sample')

if (params$save_files){
ggsave(paste(here('output/Data_exploration/'),'totalcounts.pdf', sep=''))
}

ggplot(data=cells %>% group_by(ImageNumber) %>% count(meta_description)) + geom_boxplot(aes(x=meta_description, y=n)) + theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle('Cell type counts per sample') + ylim(0,100)

if (params$save_files){
ggsave(paste(here('output/Data_exploration/'),'samplecounts.pdf', sep=''))
}


cell_occurences <- merge(cell_counts %>% tidyr::expand(tnumber, meta_description), cell_counts, by = c('tnumber','meta_description'), all.x = TRUE) %>%
                  mutate(n = coalesce(n, 0)) %>%
                  mutate(category = ifelse(n==0,'0', ifelse(n<5, '0-5', ifelse(n<20,'5-20',ifelse(n<50, '20-50','>50'))))) %>% group_by(meta_description) %>%
                  count(category)

lvls <- cell_occurences %>% filter(category == '0') %>% arrange(desc(n)) %>% pull(meta_description)

ggplot(cell_occurences, aes(x=factor(meta_description, level = lvls), y=n, fill=factor(category, levels=c('>50','20-50','5-20','0-5','0')))) + 
    geom_bar(position="fill", stat="identity") + theme_bw()  +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  ggtitle('Cell type counts in samples') + ylab('proportion') + xlab('')+ guides(fill=guide_legend(title="occurences"))

if (params$save_files){
ggsave(paste(here('output/Data_exploration/'),'celltypecounts.pdf', sep=''))
}

# merge1 <- c('MHC I^{hi}CD57^{+}', 'MHC^{hi}CD15^{+}', 'MHC I & II^{hi}')
# merge2 <- c('ER^{hi}CXCL12^{+}', 'CK8−18^{+} ER^{hi}')
# merge3 <- c(getTumorAndTMETypes()$tumor_types)
# merged_cell_counts <- cell_counts %>% mutate('merged_celltype' = ifelse(meta_description %in% getTumorAndTMETypes()$tumor_types, 'tumor', meta_description)) %>% group_by(tnumber, merged_celltype) %>% summarise(across(c(n), sum))
# 
# cell_occurences <- merge(merged_cell_counts %>% tidyr::expand(merged_celltype), merged_cell_counts, by = c('tnumber','merged_celltype'), all.x = TRUE) %>%
#                   mutate(n = coalesce(n, 0)) %>%
#                   mutate(category = ifelse(n==0,'0', ifelse(n<5, '0-5', ifelse(n<20,'5-20',ifelse(n<50, '20-50','>50'))))) %>% group_by(merged_celltype) %>%
#                   count(category)
# 
# lvls <- cell_occurences %>% filter(category == '0') %>% arrange(desc(n)) %>% pull(merged_celltype)
# 
# ggplot(cell_occurences, aes(x=factor(merged_celltype, level = lvls), y=n, fill=factor(category, levels=c('>50','20-50','5-20','0-5','0')))) + 
#     geom_bar(position="fill", stat="identity") + theme_bw()  +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
#   ggtitle('Cell type counts in samples') + ylab('proportion') + xlab('')+ guides(fill=guide_legend(title="occurences"))

```
### Compare figures to NABUCCO data.
Images from the METABRIC cohort are small compared to the NABUCCO data and therefore have a lower cell count.

```{r}
nabucco_data <- read_tsv(here('DATA/nabucco_1_densities.tsv'))

ggplot(data= nabucco_data %>% group_by(tnumber) %>% count(n)) + geom_bar(aes(y=n, x=tnumber), stat='identity') + theme_bw() + ggtitle('Total counts per sample in NABUCCO cohort')  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

if (params$save_files){
ggsave(paste(here('output/Data_exploration/'),'NABUCCO_totalcounts.pdf', sep=''))
}

ggplot(data=nabucco_data) + geom_boxplot(aes(x=phenotype, y=n)) + theme_bw() + ylim(0,1000) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle('Cell type counts per sample in NABUCCO cohort')

if (params$save_files){
ggsave(paste(here('output/Data_exploration/'),'NABUCCO_samplecounts.pdf', sep=''))
}

NABUCCO_cell_occurences <- merge(nabucco_data %>% expand(tnumber, phenotype), nabucco_data, by = c('tnumber','phenotype'), all.x = TRUE) %>%
                  mutate(n = coalesce(n, 0)) %>%
                  mutate(category = ifelse(n==0,'0', ifelse(n<10, '0<10', ifelse(n<50,'10<50','>50')))) %>% group_by(phenotype) %>%
                  count(category)

lvls <- NABUCCO_cell_occurences %>% filter(category == '>50') %>% arrange(n) %>% pull(phenotype)

ggplot(NABUCCO_cell_occurences, aes(x=factor(phenotype, level = lvls), y=n, fill=factor(category, levels=c('>50','10<50','0<10','0')))) + 
    geom_bar(position="fill", stat="identity") + theme_bw()  +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  ggtitle('Cell type counts in samples in NABUCCO cohort') + ylab('proportion') + xlab('')+ guides(fill=guide_legend(title="occurences"))

if (params$save_files){
ggsave(paste(here('output/Data_exploration/'),'NABUCCO_celltypecounts.pdf', sep=''))
}

```

## Tumor cell specifity
```{r}
total <- cell_counts %>% mutate(type = ifelse(meta_description %in% getTumorAndTMETypes()$tumor_types, 'epi', 'TME')) %>% group_by(tnumber, type) %>% summarise(total_n = sum(n))
total <- dcast(total, tnumber ~ type)
cell_counts <- merge(cell_counts, total, by='tnumber',all.x=T)
cell_counts <- cell_counts %>% mutate(percentage = ifelse(meta_description %in% getTumorAndTMETypes()$tumor_types, n/epi, n/TME))

ggplot(cell_counts %>% filter(meta_description == "Basal" & percentage > 0.05)) + geom_histogram(aes(x=percentage),binwidth = 0.05)

ggplot(cell_counts %>% filter(n > 0) %>% mutate(n = ifelse(n > 200, 200, n))) + geom_histogram(aes(x=n),binwidth = 5) + facet_wrap(vars(meta_description))
ggsave(here('output/Data_exploration/cellCountsHistograms.pdf'),height=15,width=20)

ggplot(cell_counts %>% filter(percentage > 0.05) %>% mutate(n = ifelse(n > 200, 200, n))) + geom_histogram(aes(x=percentage),binwidth = 0.05) + facet_wrap(vars(meta_description)) + theme_bw() + xlab('proportion epithelial or TME')
ggsave(here('output/Data_exploration/cellproportionsHistograms.pdf'),height=15,width=20)
```


## Cell type combinations
In future steps we are going to estimate distances between cell type combinations. Not every combination is present in every sample however.
We have a large sample size and this possibly compensates for the low cell type count. Cell types that are too rare cannot be considered.

```{r}
print(paste('number of samples that cannot be estimated: ', nrow(cell_occurences_combinations %>% filter(n_from == 0 | n_to ==0))))

heatmap_occurences <- function(dataset, occ_number){
  cell_occurences_categories <- dataset %>% mutate(category = ifelse(n_to > occ_number & n_from> occ_number, paste(as.character(occ_number),'+',sep=''), as.character(occ_number))) %>%
                                                                       group_by(phenotype_from, phenotype_to) %>%
                                                                                      count(category) %>% mutate(n = n/792)
  
  df <- cell_occurences_categories %>% filter(category == paste(as.character(occ_number),'+',sep=''))

  trans_df <- xtabs(n ~ phenotype_from + phenotype_to, df)
  hm <- Heatmap(trans_df, cluster_rows = T, cluster_columns = T, row_names_gp = gpar(fontsize=6), column_names_gp = gpar(fontsize=6),
                name = paste('sample percentage \n with', paste(as.character(occ_number),'+',sep=''), 'occ.'), column_title = 'phenotype_to', row_title = 'phenotype_from')
  
  print(hm)
  
  if (params$save_files){
  save_pdf(hm, paste(here('output/Data_exploration/'), occ_number, 'plus_occurences_heatmap.pdf', sep=''), width=6, height=4)
}
}

heatmap_occurences(cell_occurences_combinations, 0)
heatmap_occurences(cell_occurences_combinations, 5)
heatmap_occurences(cell_occurences_combinations, 10)
heatmap_occurences(cell_occurences_combinations, 20)

```

Look in PAM50 subtypes.

```{r}
cell_occurences_combinations <- getCombinationCounts()

heatmap_occurences <- function(dataset, occ_number, subtype){
  cell_occurences_categories <- dataset %>% mutate(category = ifelse(n_to > occ_number & n_from> occ_number, paste(as.character(occ_number),'+',sep=''), as.character(occ_number))) %>%
                                                                       group_by(phenotype_from, phenotype_to) %>%
                                                                                      count(category) %>% 
                                                                      mutate(n = n/ length(unique(dataset %>% pull(tnumber))))
  
  df <- cell_occurences_categories %>% filter(category == paste(as.character(occ_number),'+',sep=''))

  trans_df <- xtabs(n ~ phenotype_from + phenotype_to, df)
  hm <- Heatmap(trans_df, cluster_rows = F, cluster_columns = F, row_names_gp = gpar(fontsize=6), column_names_gp = gpar(fontsize=6),
                name = paste('sample percentage \n of type', subtype, '\n with', paste(as.character(occ_number),'+',sep=''), 'occ.'), column_title = 'phenotype_to', row_title = 'phenotype_from')
  
  print(hm)
}

heatmap_occurences(cell_occurences_combinations, 0,'ALL')
subtypes = group_split(clinical_data %>% group_by(PAM50))
for (s in 1:length(subtypes)){
  subtype = subtypes[[s]]
  heatmap_occurences(cell_occurences_combinations %>% filter(tnumber %in% subtype$ImageNumber), 0,unique(subtype$PAM50))
}

```

If we remove combinations in samples with a threshold count, the number of samples decreases fast.

```{r}
thresholds <- tibble(threshold = c(0:200))
                     
compute_from <- function(t){
   return(nrow(cell_occurences_combinations %>% filter(n_to >= t)))
}

compute_both <- function(t){
   return(nrow(cell_occurences_combinations %>% filter(n_to >= t & n_from >= t)))
}

n_from <- apply(thresholds,1, compute_from)
n_both <- apply(thresholds,1, compute_both)

thresholds <- cbind(thresholds, n_from, n_both)

ggplot(data=thresholds) + geom_point(aes(x=threshold, y=n_from, color = 'phenotype_from OR phenotype_to'))  + 
  geom_point(aes(x=threshold, y=n_both, color='phenotype_from AND phenotype_to')) + theme_bw() + ggtitle('Number of samples with at least x cells') + ylab('count') + xlab('n cells') + xlim(0,50)

if (params$save_files){
ggsave(here('output/Data_exploration/occurence_slope.pdf'))
}

```

Since there are many cell types, not all combinations are present in all slides. Some combinations have to be analysed in a subset of the images.

```{r}
cell_occurences_categories <- cell_occurences_combinations %>% mutate(category = ifelse(n_to > 5 & n_from> 20, 'from20to5', 'fewer')) %>%
                                                                     group_by(phenotype_from, phenotype_to) %>% dplyr::count(category) %>% mutate(n = n/794)

df <- cell_occurences_categories %>% filter(category == 'from20to5')

trans_df <- xtabs(n ~ phenotype_from + phenotype_to, df)
hm <- Heatmap(trans_df, cluster_rows = T, cluster_columns = T, row_names_gp = gpar(fontsize=6), column_names_gp = gpar(fontsize=6),
              name = paste('sample percentage \n with', 'from20to5', 'occ.'), column_title = 'phenotype_to', row_title = 'phenotype_from')

print(hm)

save_pdf(hm, paste(here('output/Data_exploration/'), 'from20to5', 'plus_occurences_heatmap.pdf', sep=''), width=6, height=4)

```



## Clinical data
Look at the distribution of breast cancer subtypes.

```{r}
clinical_data <- read_csv(here('DATA/IMCClinical.csv'))
clinical_data_ext <- read_tsv(here('DATA/brca_metabric_clinical_data.tsv')) %>% select(c(`Patient ID`, `ER Status`, `ER status measured by IHC`,`HER2 Status`, `PR Status`, `3-Gene classifier subtype`)) #https://www.cbioportal.org/study/clinicalData?id=brca_metabric
#https://pubmed-ncbi-nlm-nih-gov.tudelft.idm.oclc.org/27161491/
  
clinical_data_Rueda <- read.delim(here('DATA/41586_2019_1007_MOESM8_ESM.txt')) %>% select(c(`METABRIC.ID`, `ER.Expr`, `PR.Expr`, `Her2.Expr`)) #https://www-nature-com.tudelft.idm.oclc.org/articles/s41586-019-1007-8#MOESM7

all_clinical_data <- merge(merge(merge(tibble(metabric_id = unique(cells$metabric_id)), clinical_data, by='metabric_id', all.x=T), clinical_data_ext, by.x = 'metabric_id', by.y='Patient ID',all.x=T), clinical_data_Rueda, by.x='metabric_id', by.y='METABRIC.ID',all.x=T)

contingency_table <- function(feature1, feature2){
  clinical_data <- merge(all_clinical_data %>% dplyr::select(c('metabric_id', feature1)),
                         all_clinical_data %>% dplyr::select(c('metabric_id', feature2)), by='metabric_id',all.x=T)

  table(clinical_data[,2], clinical_data[,3])
  
}

order <- tibble(intclust = c("IntClust 1", "IntClust 2",  "IntClust 3",  "IntClust 4-", "IntClust 4+", "IntClust 5-", "IntClust 5+", "IntClust 6",  "IntClust 7", "IntClust 8", "IntClust 9",  "IntClust 10"))

# Contingency tables
contingency_table('HER2 Status',"Her2.Expr" )
contingency_table('ER Status',"ER.Expr" )
contingency_table('PR Status',"PR.Expr" )
# These columns are the same so can be removed
all_clinical_data <- all_clinical_data %>% select(-c(`ER.Expr`, `PR.Expr`, `Her2.Expr`))
contingency_table('ERStatus', 'ER status measured by IHC')
# These columns are also the same
all_clinical_data <- all_clinical_data %>% select(-c(`ERStatus`))
contingency_table("3-Gene classifier subtype", 'HER2 Status')
contingency_table('PAM50', 'ER Status')
contingency_table('PAM50', 'HER2 Status')
contingency_table('PAM50', 'PR Status')

all_clinical_data <- all_clinical_data %>% mutate(ER_HER2_status = ifelse(`ER Status` == 'Positive' & `HER2 Status` =='Positive', 'ER+HER2+', ifelse(`ER Status` == 'Positive' & `HER2 Status` =='Negative', 'ER+HER2-', ifelse(`ER Status` == 'Negative' &`HER2 Status` == 'Positive', 'ER-HER2+', 'ER-HER2-'))))

write_csv(all_clinical_data, here('DATA/ClinicalMerged.csv'))


ggplot(all_clinical_data) + geom_bar(aes(x=PAM50, fill=`ER_HER2_status`)) + theme_bw() + ggtitle('ER and HER2 status in PAM50 subtypes')
ggsave(here('output/Data_exploration/ERHER2statusInPAM50.pdf'),width=8,heigh=6)

# subtypes in data
ggplot(all_clinical_data) + geom_bar(aes(x=PAM50)) + theme_bw() + ggtitle('PAM50 subtype occurences')
ggsave(here('output/Data_exploration/PAM50subtypes.pdf'),width=8,heigh=6)

```

## Slides
Look at sampled images.

```{r}

TME =  c("B cells","CD38^{+} lymphocytes","CD4^{+} T cells","CD4^{+} T cells & APCs","CD57^{+}","CD8^{+} T cells","Endothelial","Fibroblasts","Fibroblasts FSP1^{+}","Granulocytes", "Ki67^{+}","Macrophages","Macrophages & granulocytes", "Myofibroblasts","Myofibroblasts PDPN^{+}", "T_{Reg} & T_{Ex}")
tumor = setdiff(unique(cells %>% pull(meta_description)), TME)

samples <- sample(unique(cells$ImageNumber),5)

for (s in samples){
  print(show_slide(s, tumor))
}

i = 2
print(show_slide(i,tumor))
ggsave(paste('output/Slides/exploreslide_',i,'.pdf',sep=''), width = 15, height=10)
```

```{r}
img_number <- cell_counts %>% filter(meta_description == "Ep CD57^{+}" & n >25)  %>% pull(tnumber)
show_slide(sample(img_number,1),"Ep CD57^{+}" )
```
## Slide heterogeneity
```{r}
print(paste('All images:', length(unique(cells %>% pull(ImageNumber))), 'of', length(unique(cells %>% pull(metabric_id))), 'patients'))
print(paste('Images with tumor tissue:', length(unique(cells %>% filter(isTumour==T) %>% pull(ImageNumber))), 'of', length(unique(cells %>% filter(isTumour==T)  %>% pull(metabric_id))), 'patients'))
print(paste('Images from normal tissue:', length(unique(cells %>% filter(is_normal == 1) %>% pull(ImageNumber))), 'of', length(unique(cells %>% filter(is_normal == 1) %>% pull(metabric_id))), 'patients'))
print(paste('in situ carcinoma tissue:', length(unique(cells %>% filter(is_dcis == 1) %>% pull(ImageNumber))), 'of', length(unique(cells %>% filter(is_dcis == 1) %>% pull(metabric_id))), 'patients'))

```


## Expression data
Cell types are classified based on expression profiles.

```{r}
library(ComplexHeatmap)
expression_profiles_summary <- as.data.frame(cells %>% group_by(meta_description)  %>% 
  summarise_at(vars("Histone H3":"DNA2"), mean))
rownames(expression_profiles_summary) <- expression_profiles_summary$meta_description
expression_profiles_summary <- scale(subset(expression_profiles_summary, select=-c(meta_description)))

expression_profiles_summary <- expression_profiles_summary[c(getTumorAndTMETypes()$tumor_types, getTumorAndTMETypes()$TME_types), ]

rows <- rowAnnotation(type = c(rep('Epithelial',16), rep('TME',16)))
  
hm <- Heatmap(expression_profiles_summary, name = 'mean expression', column_title = 'protein', row_title = 'cell type',
        row_names_gp = gpar(fontsize=6), column_names_gp = gpar(fontsize=6), cluster_rows = F, cluster_columns =  F, right_annotation  = rows)

print(hm)

save_pdf(hm, here('output/Data_exploration/expression_profiles.pdf'), width=6, height=4)
```

```{r}
cellsAlternative <- getCellsAlternative()

expression_profiles_summary <- as.data.frame(cellsAlternative %>% group_by(meta_description)  %>% 
  summarise_at(vars("Histone H3":"DNA2"), mean))
rownames(expression_profiles_summary) <- expression_profiles_summary$meta_description
expression_profiles_summary <- scale(subset(expression_profiles_summary, select=-c(meta_description)))

expression_profiles_summary <- expression_profiles_summary[c(c("CK+_ERorHER2+_cells","CK+_ERorHER2-_cells","CK-MHC+_cells","CK-_cells", "CKmed_ERorHER2-_cells"), setdiff(rownames(expression_profiles_summary),c("CK+_ERorHER2+_cells","CK+_ERorHER2-_cells","CK-MHC+_cells","CK-_cells", "CKmed_ERorHER2-_cells"))), ]

rows <- rowAnnotation(type = c(rep('epithelial',5), rep('TME',10)))
  
hm <- Heatmap(expression_profiles_summary, name = 'mean expression', column_title = 'protein', row_title = 'cell type',
        row_names_gp = gpar(fontsize=6), column_names_gp = gpar(fontsize=6), cluster_rows = F, cluster_columns =  F, right_annotation  = rows)

print(hm)

save_pdf(hm, here('output/Data_exploration/expression_profiles_alternative.pdf'), width=6, height=4)
```


