---
title: "density result exploration"
author: "Niek Brouwer"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: "hide"
params:
  save_files:
    value: FALSE
    choices:
      - FALSE
      - TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Density feature exploration

This document analyses the density features as computed in the density_method_implementation.Rmd file.

```{r}
library(tidyverse)
library(ggplot2)
library(data.table)
library(parallel)
library(here)
library(fst)
library(assertthat)
library(RColorBrewer)
library(cowplot)
library(ggpubr)
library(gtools)
source(here("UtilityFunctions.R"))
source(here("MEP_UtilityFunctions.R"))
library('glmnet', quiet = T)
library('ROCR', quiet = T)
library(ComplexHeatmap)
library(factoextra)
library(corrplot)
library(tidyHeatmap)
library(circlize)


# set seed
projectSeed <- 89230689
set.seed(projectSeed)

clinical_data <- getClinical()

```
We are going to analyse the feature sets on 3 levels: Cohort-wide, molecular subtypes and PAM50 subtypes.

## Heatmaps

### Cohort-wide
We scale the features. Rows are already scaled where the sum of all epithelial cell types == 1 and all TME cell types == 1.
```{r}
cells <- getCells()
cellPhenotypes <- getCellProportionsPerImage()
density_matrix <- generate_matrix(cellPhenotypes, 'ImageNumber')

library(circlize)
col_fun = colorRamp2(c(-4, 0, 4), c("blue", "white", "red"))

row_ha = rowAnnotation(isTumour = density_matrix %>% pull('isTumour'))

density_hm <- Heatmap(as.matrix(density_matrix %>% select(-c('isTumour')) %>% rename_with(~str_remove(., '_CPh'))),name='densities', column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=8),cluster_rows = T, cluster_columns = T, col=col_fun, right_annotation = row_ha)

save_pdf(density_hm, here('output/Method_comparison/heatmaps/density_heatmap.pdf'), width=10, height=10)

```

```{r}
cells <- getCellsAlternative()
cellPhenotypes <- getCellProportionsPerImageAlternative()
density_matrixAlternative <- generate_matrix(cellPhenotypes, 'ImageNumber')

library(circlize)
col_fun = colorRamp2(c(-4, 0,4 ), c("blue","white", "red"))

row_ha = rowAnnotation(isTumour = density_matrixAlternative %>% pull('isTumour'))

density_hm <- Heatmap(as.matrix(density_matrixAlternative %>% select(-c('isTumour')) %>% rename_with(~str_remove(., '_CPh'))),name='densities', column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=8),cluster_rows = T, cluster_columns = T, col=col_fun, right_annotation = row_ha)

save_pdf(density_hm, here('output/Method_comparison/heatmaps/density_heatmap_alternative.pdf'), width=10, height=10)
```

Take a look at the molecular subtypes.

```{r}
subtypes <- setdiff(unique(merge(density_matrix, clinical_data, by.x='row.names', by.y='ImageNumber', all.x=T) %>% pull("ER_HER2_status")),NA)

for (s in subtypes){
  samples_cellPHenotypes <- density_matrix %>% filter(rownames(density_matrix) %in% (clinical_data %>% filter(`ER_HER2_status` == s) %>% pull(`ImageNumber`)))
  
  row_ha = rowAnnotation(isTumour = samples_cellPHenotypes %>% pull('isTumour'))

  subtype_hm <- Heatmap(as.matrix(samples_cellPHenotypes %>% select(-c('isTumour'))  %>% rename_with(~str_remove(., '_CPh'))),name=paste(s, 'samples'), column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=8),cluster_rows = T, cluster_columns = T, col=col_fun, right_annotation = row_ha)
  
  save_pdf(subtype_hm, paste(here('output/Method_comparison/heatmaps/molecularSubtype_heatmaps/'),'density_heatmap_',gsub('+', 'pos', s,fixed=T),'.pdf',sep=''), width=10, height=10)
  
  
  samples_cellPHenotypesAlternative <- density_matrixAlternative %>% filter(rownames(density_matrix) %in% (clinical_data %>% filter(`ER_HER2_status` == s) %>% pull(`ImageNumber`)))

    row_ha = rowAnnotation(isTumour = samples_cellPHenotypesAlternative %>% pull('isTumour'))

  subtype_hm <- Heatmap(as.matrix(samples_cellPHenotypesAlternative %>% select(-c('isTumour'))  %>% rename_with(~str_remove(., '_CPh'))),name=paste(s, 'samples'), column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=8),cluster_rows = T, cluster_columns = T, col=col_fun, right_annotation = row_ha)
  
  save_pdf(subtype_hm, paste(here('output/Method_comparison/heatmaps/molecularSubtype_heatmaps/'),'density_heatmap_',gsub('+', 'pos', s,fixed = T),'_alternative.pdf',sep=''), width=10, height=10)

  
}
```


Take a look at the PAM50 patient groups.

```{r}
subtypes <- setdiff(unique(merge(density_matrix, clinical_data, by.x='row.names', by.y='ImageNumber', all.x=T) %>% pull('PAM50')),NA)

for (s in subtypes){
  samples_cellPHenotypes <- density_matrix %>% filter(rownames(density_matrix) %in% (clinical_data %>% filter(`PAM50` == s) %>% pull(`ImageNumber`)))
  
  row_ha = rowAnnotation(isTumour = samples_cellPHenotypes %>% pull('isTumour'))

  subtype_hm <- Heatmap(as.matrix(samples_cellPHenotypes %>% select(-c('isTumour'))  %>% rename_with(~str_remove(., '_CPh'))),name=paste(s, 'samples'), column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=8),cluster_rows = T, cluster_columns = T, col=col_fun, right_annotation = row_ha)
  
  save_pdf(subtype_hm, paste(here('output/Method_comparison/heatmaps/PAM50_heatmaps/'),'density_heatmap_',s,'.pdf',sep=''), width=10, height=10)
  
  
  samples_cellPHenotypesAlternative <- density_matrixAlternative %>% filter(rownames(density_matrix) %in% (clinical_data %>% filter(`PAM50` == s) %>% pull(`ImageNumber`)))

    row_ha = rowAnnotation(isTumour = samples_cellPHenotypesAlternative %>% pull('isTumour'))

  subtype_hm <- Heatmap(as.matrix(samples_cellPHenotypesAlternative %>% select(-c('isTumour'))  %>% rename_with(~str_remove(., '_CPh'))),name=paste(s, 'samples'), column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=8),cluster_rows = T, cluster_columns = T, col=col_fun, right_annotation = row_ha)
  
  save_pdf(subtype_hm, paste(here('output/Method_comparison/heatmaps/PAM50_heatmaps/'),'density_heatmap_',s,'_alternative.pdf',sep=''), width=10, height=10)

  
}
```


## PCA
Now perform PCA to separate the samples.

```{r}
generate_labels <- function(matrix){
labels <- merge(x = as_tibble(rownames(matrix)), y = clinical_data, by.x='value', by.y='ImageNumber', all.x=T)
labels <- labels %>% rename(ImageNumber = value)
# %>%
#   mutate(HER = ifelse(PAM50=='Her2', 'HER2', 'Other')) %>%
#   mutate(LuminalA = ifelse(PAM50=='LumA', 'Luminal A', 'Other')) %>%
#   mutate(LuminalB = ifelse(PAM50=='LumB', 'Luminal B', 'Other')) %>%
#   mutate(Bas = ifelse(PAM50=='Basal', 'Basal', 'Other')) %>%
#   mutate(Norm= ifelse(PAM50=='Normal', 'Normal', 'Other'))

colnames(labels) <- gsub(" ", "_", colnames(labels))
return(labels)
}


plot_pca <- function(matrix, label){
  labels = generate_labels(matrix)
  matrix_withlabels <- cbind.data.frame(matrix, rownames(matrix))
  colnames(matrix_withlabels)[ncol(matrix_withlabels)] = 'ImageNumber'
  matrix_withlabels <- merge(matrix_withlabels, labels, by='ImageNumber', all.x=T)

  res.pca <- prcomp(matrix_withlabels %>% select(-colnames(labels)), scale = FALSE)
  print(fviz_eig(res.pca))
  print(fviz_pca_biplot(res.pca, label="var", habillage=matrix_withlabels %>% pull(!!label), addEllipses = T, ellipse.level = 0.95,select.var = list(contrib = 5) ))
  
}

plot_pca(density_matrix %>% filter(isTumour > 0) %>% select(-c(isTumour)), 'PAM50')
ggsave(here('output/Method_comparison/pca/pca_density_PAM50.pdf'),width=8,height=7)

plot_pca(density_matrixAlternative  %>% filter(isTumour > 0) %>% select(-c(isTumour)), 'PAM50')
ggsave(here('output/Method_comparison/pca/pca_density_PAM50_alternative.pdf'),width=8,height=7)

plot_pca(density_matrix  %>% filter(isTumour > 0) %>% select(-c(isTumour)), "ER_HER2_status")
ggsave(here('output/Method_comparison/pca/pca_density_ERHer2.pdf'),width=8,height=7)

plot_pca(density_matrixAlternative  %>% filter(isTumour > 0) %>% select(-c(isTumour)),"ER_HER2_status")
ggsave(here('output/Method_comparison/pca/pca_density_ERHer2_alternative.pdf'),width=8,height=7)
```





## Basal subtypes with ER+ cells

Patients with basal subtypes contain ER+ cells. Lets look at the slides of these patients.

```{r}
positive_celltypes <- c("ER^{hi}CXCL12^{+}", "CK8-18^{+} ER^{hi}")

Basal_cellPHenotypes <- tumor_densities %>% filter(rownames(tumor_density_matrix) %in% (clinical_data %>% filter(`PAM50` == 'Basal') %>% pull(`ImageNumber`)))
positive_samples <- Basal_cellPHenotypes %>% filter(get("CK8-18^{+} ER^{hi}_CPh") > 0 | get("ER^{hi}CXCL12^{+}_CPh") > 0)  %>% pull('ImageNumber')

# count_features <- getDensityFeatures()
# counts_positive_samples <- count_features %>% filter(ImageNumber %in% (positive_samples %>% pull('ImageNumber')))

for (j in 1:10){
  i = positive_samples[j]
  s <- show_slide(i,positive_celltypes)
  print(s)
  ggsave(paste('output/Slides/basal_slide',i,'.pdf',sep=''), width = 15, height=10)
}

```
In some samples the ER+ cell count is very low.

```{r}
basal_counts <- cells %>% dplyr::count(ImageNumber, meta_description) %>% mutate(type = ifelse((ImageNumber %in% positive_samples),'Basal with ER+ cells','Other types with ER+ cells'))

basal_counts <- merge(basal_counts, clinical_data %>% select(c("ImageNumber", "PAM50")), by='ImageNumber')

ggplot(data=basal_counts %>% filter(meta_description %in% positive_celltypes)) + theme_bw() + geom_boxplot(aes(x=PAM50,y=n,color=meta_description)) + ylim(0,50) + ggtitle('ER+ cell count in samples marked as basal and other types ')

ggsave(here('output/Density_results/ERcellCount_Basalsubtype.pdf'))

```


Look at the expression of these cells

```{r}
positive_basal_cells <- cells %>% filter(meta_description %in% positive_celltypes) %>% mutate(type = ifelse((ImageNumber %in% positive_samples),'Basal with ER+ cells','Other types with ER+ cells'))
ggplot(data=positive_basal_cells) + theme_bw() + xlab('ER expression') + geom_boxplot(aes(x=type, y=ER)) + ylim(0,200) + ggtitle('Meassured ER expression of ER+ cells in samples marked as Basal and other type') + xlab('sample type') + ylab('ER expression')

ggsave(here('output/Density_results/ERexpression_Basalsubtype.pdf'))

```

The expression of these cells is not lower as other ER+ cells and are therefore correctly marked as ER+ cells.

```{r}
subtypes <- setdiff(unique(clinical_data %>% pull(PAM50)), NA)
cell_count_matrix <- getDensityFeatures() %>% select(-c('isTumour'))
cell_count_matrix_withlabels <- merge(cell_count_matrix, clinical_data %>% select(c(PAM50, ImageNumber)), by='ImageNumber', all.x=T)

zero_count_result <- tibble(features = colnames(cell_count_matrix)[2:33])
for (s in 1:length(subtypes)){
  subset <- t(cell_count_matrix_withlabels %>% filter(`PAM50` == subtypes[[s]]))
  subset <- subset[2:33,]
  subset_m <- matrix(as.numeric(subset),    # Convert to numeric matrix
                    ncol = ncol(subset))
  zero_count <- rowSums(subset_m != 0)  / ncol(subset_m)
  zero_count_result <- cbind(zero_count_result, zero_count)
  
}

colnames(zero_count_result) <- c('feature', subtypes)
rownames(zero_count_result) <- zero_count_result$feature
zero_count_result <- zero_count_result[,-1]

h <- Heatmap(zero_count_result,name='not zero percentage', cluster_columns = F, cluster_rows = T)
save_pdf(h, here('output/Density_results/zero_counts_per_subtype.pdf'),width=10, height=10)

```

## Densities in PAM50 subtypes

```{r}
library(matrixStats)

count_result <- tibble(features = colnames(cell_count_matrix)[2:33])
for (s in 1:length(subtypes)){
  subset <- t(cell_count_matrix_withlabels %>% filter(`PAM50` == subtypes[[s]]))
  subset <- subset[2:33,]
  subset_m <- matrix(as.numeric(subset),    # Convert to numeric matrix
                    ncol = ncol(subset))
  means <- rowMeans(subset_m)
  count_result <- cbind(count_result, means)
  
}

colnames(count_result) <- c('feature', subtypes)
rownames(count_result) <- count_result$feature
count_result <- count_result[,-1]

h <- Heatmap(count_result,name='mean count', cluster_columns = F, cluster_rows = F)
save_pdf(h, here('output/Density_results/mean_count_per_subtype.pdf'),width=10, height=10)

count_result <- tibble(features = colnames(cell_count_matrix)[2:33])
for (s in 1:length(subtypes)){
  subset <- t(cell_count_matrix_withlabels %>% filter(`PAM50` == subtypes[[s]]))
  subset <- subset[2:33,]
  subset_m <- matrix(as.numeric(subset),    # Convert to numeric matrix
                    ncol = ncol(subset))
  means <- rowMedians(subset_m)
  count_result <- cbind(count_result, means)
  
}

colnames(count_result) <- c('feature', subtypes)
rownames(count_result) <- count_result$feature
count_result <- count_result[,-1]

h <- Heatmap(count_result,name='median count', cluster_columns = F, cluster_rows = F)
save_pdf(h, here('output/Density_results/median_count_per_subtype.pdf'),width=10, height=10)


```
```{r}
library(matrixStats)

cellPhenotypes_withlabels <- merge(x=cellPhenotypes, y= clinical_data %>% select(c('ImageNumber', 'PAM50')), by='ImageNumber', all.x=T)
cellPhenotypes_withlabels <- na.omit(cellPhenotypes_withlabels)

count_result <- tibble(features = colnames(cellPhenotypes_withlabels)[2:33])
for (s in 1:length(subtypes)){
  subset <- cellPhenotypes_withlabels %>% filter(`PAM50` == subtypes[[s]])
  means <- colMeans(subset[,2:33])
  count_result <- cbind(count_result, means)
  
}

colnames(count_result) <- c('feature', subtypes)
rownames(count_result) <- count_result$feature
count_result <- count_result[,-1]

h <- Heatmap(count_result,name='mean proportion', cluster_columns = F, cluster_rows = T)
save_pdf(h, here('output/Density_results/mean_proportion_per_subtype.pdf'),width=10, height=10)

count_result <- tibble(features = colnames(cellPhenotypes_withlabels)[2:33])
for (s in 1:length(subtypes)){
  subset <- cellPhenotypes_withlabels %>% filter(`PAM50` == subtypes[[s]])
  means <- apply(subset[,2:33],2,median)
  count_result <- cbind(count_result, means)
  
}

colnames(count_result) <- c('feature', subtypes)
rownames(count_result) <- count_result$feature
count_result <- count_result[,-1]

h <- Heatmap(count_result,name='median proportion', cluster_columns = F, cluster_rows = T)
save_pdf(h, here('output/Density_results/median_proportion_per_subtype.pdf'),width=10, height=10)
```
