---
title: "distance_feature_analysis"
author: "Niek Brouwer"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: "hide"
params:
  save_files:
    value: FALSE
    choices:
      - FALSE
      - TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Density feature exploration
This document analyses the distance features as computed in the distance_method_implementation.Rmd file.


```{r}
library(here)
library(ggplot2)
library(tidyverse)
library(fst)
source(here("UtilityFunctions.R"))
source(here("MEP_UtilityFunctions.R"))
library(factoextra)
library(corrplot)
library(gridExtra)
library(cowplot)
library(grid)
library(ComplexHeatmap)
library(tidyHeatmap)
library(umap)
library(circlize)
library(dplyr)
library(ggsignif)
library(ggpubr)
library(matrixStats)
library(data.table)

clinical_data <- getClinical()
```

## feature estimation percentages

Compute feature estimation percentages for filter purposes.
Order features based on the max value, to maintain features that are specific for a single subtype.

```{r}
shape_matrix <- getShapeFeatures()
sub_clinical_data <- clinical_data %>% dplyr::select(c(`ImageNumber`,`PAM50`))

shape_matrix_withlabels <- merge(shape_matrix, sub_clinical_data, by.x = 'tnumber', by.y='ImageNumber', all.x=T)
subtypes <- setdiff(unique(shape_matrix_withlabels %>% pull(`PAM50`)), c(NA))

estimations <- data.frame(matrix(ncol=0, nrow=ncol(shape_matrix_withlabels)))
for (s in 1:length(subtypes)){
  subset <- shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]])
  na_count <-sapply(shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]]), function(y) sum(length(which(is.na(y)))))
  estimation_count <-  1- (na_count / nrow(shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]])))
  estimation_count <- data.frame(estimation_count)
  estimations <- cbind(estimations, estimation_count)

  }
estimations <- setNames(estimations, subtypes)
estimations <- estimations[!(row.names(estimations) %in% c("tnumber","PAM50")),]
estimations <- estimations%>% mutate(row_wise_max = rowMaxs(as.matrix(estimations)))
estimations <- estimations %>% filter(row_wise_max != 0)
estimations <- estimations %>% arrange(desc(row_wise_max)) %>% mutate(rank = row_number()) %>% mutate(rank_percentage = 100* (rank / nrow(estimations)))

saveRDS(estimations, file=here('scratch/estimationCounts.rds'))

```

## Heatmap
What patterns in the parameters do we find?

```{r}
col_fun = colorRamp2(c(-4, 0, 4), c("blue", "white", "red"))

generate_hm <- function(m, hm_name, n_clust, cluster_rows, cluster_columns,fontsize_rows, fontsize_columns, row_title, column_title){
  ht <- Heatmap(as.matrix(m), name = paste(hm_name,'features'), column_title = column_title, row_title = row_title, 
                row_names_gp = gpar(fontsize=fontsize_rows), column_names_gp = gpar(fontsize=fontsize_columns), 
                cluster_rows = cluster_rows, cluster_columns =  cluster_columns, row_km = n_clust,col=col_fun)

  return(ht)
}


subtype_proportions <- function(clusters,label){
  clinical_data <- clinical_data %>% mutate(cluster = NA)
  for (i in 1:length(clusters)){
    clinical_data <- clinical_data %>% mutate(cluster = ifelse(ImageNumber %in% clusters[[i]], as.character(names(clusters)[[i]]),cluster))
  }
  clinical_data <- clinical_data %>% drop_na(cluster)

  lengths <- tibble(cluster = names(clusters), length = c(sapply(clusters, length)))

  p <- ggplot() +
    geom_bar(data=clinical_data, aes(x= 'ALL', y= length(data),fill=get(label)),stat='identity',position='fill')  +
    geom_text(data=lengths, aes(x='ALL',y=1,label = sum(lengths$length)), vjust = -1, size=3) +
    geom_bar(data=clinical_data,aes(x= cluster, y= length(data),fill=get(label)),stat='identity',position='fill') +
    geom_text(data=lengths, aes(x=cluster,y=1,label = length), vjust = -1, size=3) +
    xlab('cluster') + ylab('proportion') + theme_bw() +ylim(0,1.2)
  return(p)
}

generate_labels <- function(matrix){
  labels <- merge(as_tibble(rownames(matrix)), clinical_data, by.x='value', by.y='ImageNumber', all.x=T) %>% 
    rename(tnumber = value) %>% mutate(tnumber = as.numeric(tnumber))
  labels <- labels %>%
    mutate(HER = ifelse(PAM50=='Her2', 'HER2', 'Other')) %>%
    mutate(LuminalA = ifelse(PAM50=='LumA', 'Luminal A', 'Other')) %>%
    mutate(LuminalB = ifelse(PAM50=='LumB', 'Luminal B', 'Other')) %>%
    mutate(Bas = ifelse(PAM50=='Basal', 'Basal', 'Other')) %>%
    mutate(Norm= ifelse(PAM50=='Normal', 'Normal', 'Other'))

colnames(labels) <- gsub(" ", "_", colnames(labels))
 return(labels) 
}

```

Generate heatmaps of shape and scale features.

Some features are rarely estimated, these should be filtered. Initially we filtered out features that occured in <50% of the samples.  This favors features in luminal A and B cells because these groups are large. Therefore we look at the estimations per subtype.

```{r}
shape_matrix <- getShapeFeatures()
sub_clinical_data <- clinical_data %>% dplyr::select(c(`ImageNumber`,`PAM50`))

shape_matrix_withlabels <- merge(shape_matrix, sub_clinical_data, by.x = 'tnumber', by.y='ImageNumber', all.x=T)
subtypes <- setdiff(unique(shape_matrix_withlabels %>% pull(`PAM50`)), c(NA))

result <- tibble(features = colnames(shape_matrix_withlabels))
for (s in 1:length(subtypes)){
  subset <- shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]])
  na_count <-sapply(shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]]), function(y) sum(length(which(!is.na(y)))))
  na_count <- as.numeric(na_count / nrow(shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]])))
  na_count <- data.table(na_count)
  result <- cbind(result, na_count)
}

colnames(result) <- c('feature', subtypes)
result <- head(result, - 1)
result <- result[-1,]

result <- result[order(result$feature),]

rownames(result) <- result$feature
result <- result[,-1]

h <- Heatmap(result,name='estimated percentage', cluster_columns = F, cluster_rows = F,row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=15))
save_pdf(h, here('output/Distance_results/Missing_estimations_analysis/not_estimated_per_subtype.pdf'),width=5, height=40)
```


```{r}
shape_matrix <- generate_matrix(getShapeFeatures(),'tnumber',NULL,0.5, scale_rows = TRUE)
scale_matrix <- generate_matrix(getScaleFeatures(),'tnumber',NULL,0.5, scale_rows = TRUE)

shape_hm <- generate_hm(shape_matrix,'shape',20, TRUE, TRUE, 2,8,'samples', 'combinations')
scale_hm <- generate_hm(scale_matrix,'scale',20, TRUE, TRUE, 2,8,'samples', 'combinations')

save_pdf(shape_hm, here('output/Method_comparison/heatmaps/shape_heatmap.pdf'), width=20, height=12)
save_pdf(scale_hm, here('output/Method_comparison/heatmaps/scale_heatmap.pdf'), width=20, height=12)

ht = draw(shape_hm)
cl = row_order(ht)
print(subtype_proportions(cl,'PAM50') + ggtitle('PAM50 subtype proportion in clusters'))

ggsave(here('output/Method_comparison/heatmaps/subtypes_in_clusters_shape.pdf'))

ht = draw(scale_hm)
cl = row_order(ht)
print(subtype_proportions(cl,'PAM50')  + ggtitle('PAM50 subtype proportion in clusters'))

ggsave(here('output/Method_comparison/heatmaps/subtypes_in_clusters_scale.pdf'))
```

## Heatmap per PAM50 subtype
```{r}

shape_matrix$tnumber <- rownames(shape_matrix)
scale_matrix$tnumber <- rownames(scale_matrix)
shape_matrix_withlabels <- merge(shape_matrix, clinical_data%>% select(ImageNumber, PAM50 ), by.x='tnumber', by.y='ImageNumber')
scale_matrix_withlabels <- merge(scale_matrix, clinical_data%>% select(ImageNumber, PAM50 ), by.x='tnumber', by.y='ImageNumber')
shape_matrix_withlabels <- shape_matrix_withlabels %>% select(-c(tnumber))
scale_matrix_withlabels <- scale_matrix_withlabels %>% select(-c(tnumber))

subtypes <- setdiff(unique(clinical_data %>% pull(`PAM50`)), c(NA))
shape_result <- tibble(features = colnames(shape_matrix_withlabels)[1:ncol(shape_matrix_withlabels)])
for (s in 1:length(subtypes)){
  subset <- t(shape_matrix_withlabels %>% filter(`PAM50` == subtypes[[s]]))
  subset <- subset[1:ncol(shape_matrix_withlabels),]
  subset_m <- matrix(as.numeric(subset),    # Convert to numeric matrix
                    ncol = ncol(subset))
  means <- rowMeans(subset_m)
  shape_result <- cbind(shape_result, means)
  
}

colnames(shape_result) <- c('feature', subtypes)
shape_result <- shape_result[order(shape_result$feature),]
rownames(shape_result) <- shape_result$feature
shape_result <- shape_result[,-1]

h <- Heatmap(shape_result,name='mean shape parameter', cluster_columns = F, cluster_rows = F,row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=15))
save_pdf(h, here('output/Distance_results/mean_shape_per_subtype.pdf'),width=5, height=40)

scale_result <- tibble(features = colnames(scale_matrix_withlabels)[1:ncol(scale_matrix_withlabels)])
for (s in 1:length(subtypes)){
  subset <- t(scale_matrix_withlabels %>% filter(`PAM50` == subtypes[[s]]))
  subset <- subset[1:ncol(scale_matrix_withlabels),]
  subset_m <- matrix(as.numeric(subset),    # Convert to numeric matrix
                    ncol = ncol(subset))
  means <- rowMeans(subset_m)
  scale_result <- cbind(scale_result, means)
  
}

colnames(scale_result) <- c('feature', subtypes)
scale_result <- scale_result[order(scale_result$feature),]
rownames(scale_result) <- scale_result$feature
scale_result <- scale_result[,-1]

h <- Heatmap(scale_result,name='mean scale parameter', cluster_columns = F, cluster_rows = F,row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=15))
save_pdf(h, here('output/Distance_results/mean_scale_per_subtype.pdf'),width=5, height=40)
```

## UMAP
Plot the parameters as UMAP to see if there are clusters of the same cancer subtype.

```{r}
colnames(generate_labels(shape_matrix))

generate_UMAP <- function(matrix, title){
  labels =generate_labels(matrix)
  umap <- umap(matrix)
  umap <- umap$layout %>%
  as.data.frame()%>%
  rename(UMAP1="V1",
         UMAP2="V2") %>%
  mutate(tnumber=row_number())%>%
  inner_join(labels, by='tnumber')

  return(umap %>%
  ggplot(aes(x = UMAP1, 
             y = UMAP2, 
             color = PAM50))+
  geom_point()+ theme_bw() +
  labs(x = "UMAP1",
       y = "UMAP2",
      subtitle = title))
}

print(generate_UMAP(shape_matrix,'UMAP plot of shape parameters'))
print(generate_UMAP(scale_matrix,'UMAP plot of scale parameters'))
```


## PCA
PCA of samples based on parameter features. We only consider features which are estimated in more than 50% of the samples. 

```{r}
plot_pca <- function(matrix, label){
  labels = generate_labels(matrix)
  matrix_withlabels <- cbind.data.frame(matrix, rownames(matrix))
  colnames(matrix_withlabels)[ncol(matrix_withlabels)] = 'tnumber'
  matrix_withlabels <- merge(matrix_withlabels, labels, by='tnumber', all.x=T)

  res.pca <- prcomp(matrix_withlabels %>% select(-colnames(labels)), scale = FALSE)
  print(fviz_eig(res.pca))
  p2 <- fviz_pca_biplot(res.pca, label="var", habillage=matrix_withlabels %>% pull(!!label), addEllipses = T, ellipse.level = 0.95,select.var = list(contrib = 5))
    
  print(p2)
  return(p2)
}

shape_matrix <- generate_matrix(getShapeFeatures(),'tnumber',NULL,0.5, scale_rows = TRUE)
scale_matrix <- generate_matrix(getScaleFeatures(),'tnumber',NULL,0.5, scale_rows = TRUE)

pca_shape <- plot_pca(shape_matrix, "PAM50")
ggsave(here('output/Method_comparison/pca/pca_shape.pdf'),width=10, height=10)
pca_scale <- plot_pca(scale_matrix, 'PAM50')
ggsave(here('output/Method_comparison/pca/pca_scale.pdf'),width=10, height=10)

```

There is no clear separation of cell types.
Look at Intclusts as well.

```{r}
pca_shape <- plot_pca(shape_matrix, "IntClust")
ggsave(here('output/Method_comparison/pca/pca_shape_IntClust.pdf'),width=10, height=10)
pca_scale <- plot_pca(scale_matrix, 'IntClust')
ggsave(here('output/Method_comparison/pca/pca_scale_IntClust.pdf'),width=10, height=10)
```

