---
title: "distance_feature_analysis"
author: "Niek Brouwer"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: "hide"
params:
  save_files:
    value: FALSE
    choices:
      - FALSE
      - TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Distance feature exploration
This document analyses the distance features as computed in the distance_method_implementation.Rmd file.


```{r}
library(here)
library(ggplot2)
library(tidyverse)
source(here("UtilityFunctions.R"))
source(here("MEP_UtilityFunctions.R"))
library(factoextra)
library(corrplot)
library(gridExtra)
library(cowplot)
library(grid)
library(ComplexHeatmap)
library(tidyHeatmap)
library(umap)
library(circlize)

clinical_data <- getClinical()
```

## Feature estimations & Filtering

Compute feature estimation percentages for filter purposes.
Order features based on the max value, to maintain features that are specific for a single subtype.

```{r}
shape_matrix <- getShapeFeatures()
print(paste('percentage unestimated:', 100*sum(is.na(shape_matrix))/(nrow(shape_matrix)*ncol(shape_matrix))))
sub_clinical_data <- clinical_data %>% dplyr::select(c(`ImageNumber`,`PAM50`))

shape_matrix_withlabels <- merge(shape_matrix, sub_clinical_data, by.x = 'tnumber', by.y='ImageNumber', all.x=T)
subtypes <- setdiff(unique(shape_matrix_withlabels %>% pull(`PAM50`)), c(NA))

estimations <- data.frame(matrix(ncol=0, nrow=ncol(shape_matrix_withlabels)))
for (s in 1:length(subtypes)){
  subset <- shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]])
  na_count <-sapply(shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]]), function(y) sum(length(which(is.na(y)))))
  estimation_count <-  1- (na_count / nrow(shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]])))
  estimation_count <- data.frame(estimation_count)
  estimations <- cbind(estimations, estimation_count)
}

estimations <- setNames(estimations, subtypes)
estimations <- estimations[!(row.names(estimations) %in% c("tnumber","PAM50")),]
estimations <- estimations%>% mutate(row_wise_max = rowMaxs(as.matrix(estimations)))
estimations <- estimations %>% filter(row_wise_max != 0)
estimations <- estimations %>% arrange(desc(row_wise_max)) %>% mutate(rank = row_number()) %>% mutate(rank_percentage = 100* (rank / nrow(estimations)))

saveRDS(estimations, file=here('scratch/estimationCounts.rds'))

```

```{r}
shape_matrix <- getShapeFeaturesAlternative()
print(paste('percentage unestimated:', 100*sum(is.na(shape_matrix))/(nrow(shape_matrix)*ncol(shape_matrix))))
sub_clinical_data <- clinical_data %>% dplyr::select(c(`ImageNumber`,`PAM50`))

shape_matrix_withlabels <- merge(shape_matrix, sub_clinical_data, by.x = 'tnumber', by.y='ImageNumber', all.x=T)
subtypes <- setdiff(unique(shape_matrix_withlabels %>% pull(`PAM50`)), c(NA))

estimationsAlternative <- data.frame(matrix(ncol=0, nrow=ncol(shape_matrix_withlabels)))
for (s in 1:length(subtypes)){
  subset <- shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]])
  na_count <-sapply(shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]]), function(y) sum(length(which(is.na(y)))))
  estimation_count <-  1- (na_count / nrow(shape_matrix_withlabels %>% filter(`PAM50` ==subtypes[[s]])))
  estimation_count <- data.frame(estimation_count)
  estimationsAlternative <- cbind(estimationsAlternative, estimation_count)

  }
estimationsAlternative <- setNames(estimationsAlternative, subtypes)
estimationsAlternative <- estimationsAlternative[!(row.names(estimationsAlternative) %in% c("tnumber","PAM50")),]
estimationsAlternative <- estimationsAlternative%>% mutate(row_wise_max = rowMaxs(as.matrix(estimationsAlternative)))
estimationsAlternative <- estimationsAlternative %>% filter(row_wise_max != 0)
estimationsAlternative <- estimationsAlternative %>% arrange(desc(row_wise_max)) %>% mutate(rank = row_number()) %>% mutate(rank_percentage = 100* (rank / nrow(estimationsAlternative)))

saveRDS(estimationsAlternative, file=here('scratch/estimationCountsAlternative.rds'))

```
What happens to the percentage of unestimated parameters if we remove combinations with very rare cell types.
```{r}
n = 0.1
rare_combinations <- rownames(estimations %>% filter(row_wise_max < n))
print(paste('percentage of rare combinations:',100*length(rare_combinations)/ncol(getShapeFeatures())))
shape_matrix <- getShapeFeatures() %>% select(-all_of(rare_combinations))
print(paste('percentage unestimated:', 100*sum(is.na(shape_matrix))/(nrow(shape_matrix)*ncol(shape_matrix))))


rare_combinations <- rownames(estimationsAlternative %>% filter(row_wise_max < n))
print(paste('percentage of rare combinations:', 100*length(rare_combinations)/ncol(getShapeFeaturesAlternative())))
shape_matrix <- getShapeFeaturesAlternative() %>% select(-all_of(rare_combinations))
print(paste('percentage unestimated:', 100*sum(is.na(shape_matrix))/(nrow(shape_matrix)*ncol(shape_matrix))))

```



## Heatmap

### Cohort-wide
Remove very rare combinations.

```{r}
col_fun = colorRamp2(c(-4, 0, 4), c("blue", "white", "red"))
dominant_combinations_shape <- rownames(estimations %>% filter(row_wise_max > 0.5))
dominant_combinations_scale <- rownames(estimations %>% filter(row_wise_max > 0.5))

shape_features <- getShapeFeatures() %>% select(c('tnumber', dominant_combinations_shape)) %>% rename_with(~paste(., '_shape',sep=''))
scale_features <- getScaleFeatures() %>% select(c('tnumber', dominant_combinations_scale)) %>% rename_with(~paste(., '_scale',sep=''))
distance_features <- merge(shape_features,scale_features, by.x='tnumber_shape', by.y='tnumber_scale')
distance_matrix <- generate_matrix(distance_features,'tnumber_shape')

density_hm <- Heatmap(as.matrix(distance_matrix),name='shape and scale', column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=2),cluster_rows = T, cluster_columns = T, col=col_fun)

save_pdf(density_hm, here('output/Method_comparison/heatmaps/shapeAndScale_heatmap.pdf'), width=10, height=10)

```

```{r}
dominant_combinations_shape <- rownames(estimationsAlternative %>% filter(row_wise_max > 0.5))
dominant_combinations_scale <- rownames(estimationsAlternative %>% filter(row_wise_max > 0.5))

shape_features <- getShapeFeaturesAlternative() %>% select(c('tnumber', dominant_combinations_shape)) %>% rename_with(~paste(., '_shape',sep=''))
scale_features <- getScaleFeaturesAlternative() %>% select(c('tnumber', dominant_combinations_scale)) %>% rename_with(~paste(., '_scale',sep=''))
distance_features <- merge(shape_features,scale_features, by.x='tnumber_shape', by.y='tnumber_scale')
distance_matrixAlternative <- generate_matrix(distance_features,'tnumber_shape')

density_hm <- Heatmap(as.matrix(distance_matrixAlternative),name='shape and scale', column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=2),cluster_rows = T, cluster_columns = T, col=col_fun)

save_pdf(density_hm, here('output/Method_comparison/heatmaps/shapeAndScale_heatmap_alternative.pdf'), width=10, height=10)
```


### Molecular subtypes

```{r}
subtypes <- setdiff(unique(merge(distance_matrix, clinical_data, by.x='row.names', by.y='ImageNumber', all.x=T) %>% pull("ER_HER2_status")),NA)

for (s in subtypes){
  samples_cellPHenotypes <- distance_matrix %>% filter(rownames(distance_matrix) %in% (clinical_data %>% filter(`ER_HER2_status` == s) %>% pull(`ImageNumber`)))


  subtype_hm <- Heatmap(as.matrix(samples_cellPHenotypes),name=paste(s, 'samples'), column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=3),cluster_rows = T, cluster_columns = T, col=col_fun)
  
  save_pdf(subtype_hm, paste(here('output/Method_comparison/heatmaps/molecularSubtype_heatmaps/'),'shapeAndScale_heatmap_',gsub('+', 'pos', s,fixed=T),'.pdf',sep=''), width=10, height=10)
  
  samples_cellPHenotypesAlternative <- distance_matrixAlternative %>% filter(rownames(distance_matrixAlternative) %in% (clinical_data %>% filter(`ER_HER2_status` == s) %>% pull(`ImageNumber`)))

  subtype_hm <- Heatmap(as.matrix(samples_cellPHenotypesAlternative),name=paste(s, 'samples'), column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=3),cluster_rows = T, cluster_columns = T, col=col_fun)
  
  save_pdf(subtype_hm, paste(here('output/Method_comparison/heatmaps/molecularSubtype_heatmaps/'),'shapeAndScale_heatmap_',gsub('+', 'pos', s,fixed = T),'_alternative.pdf',sep=''), width=10, height=10)

  
}
```

### PAM50 subtypes

```{r}
subtypes <- setdiff(unique(merge(distance_matrix, clinical_data, by.x='row.names', by.y='ImageNumber', all.x=T) %>% pull("PAM50")),NA)

for (s in subtypes){
  samples_cellPHenotypes <- distance_matrix %>% filter(rownames(distance_matrix) %in% (clinical_data %>% filter(`PAM50` == s) %>% pull(`ImageNumber`)))


  subtype_hm <- Heatmap(as.matrix(samples_cellPHenotypes),name=paste(s, 'samples'), column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=3),cluster_rows = T, cluster_columns = T, col=col_fun)
  
  save_pdf(subtype_hm, paste(here('output/Method_comparison/heatmaps/PAM50_heatmaps/'),'shapeAndScale_heatmap_',gsub('+', 'pos', s,fixed=T),'.pdf',sep=''), width=10, height=10)
  
  samples_cellPHenotypesAlternative <- distance_matrixAlternative %>% filter(rownames(distance_matrixAlternative) %in% (clinical_data %>% filter(`PAM50` == s) %>% pull(`ImageNumber`)))

  subtype_hm <- Heatmap(as.matrix(samples_cellPHenotypesAlternative),name=paste(s, 'samples'), column_title = 'cell type', 
                      row_title = 'sample', row_names_gp = gpar(fontsize=2), column_names_gp = gpar(fontsize=3),cluster_rows = T, cluster_columns = T, col=col_fun)
  
  save_pdf(subtype_hm, paste(here('output/Method_comparison/heatmaps/PAM50_heatmaps/'),'shapeAndScale_heatmap_',gsub('+', 'pos', s,fixed = T),'_alternative.pdf',sep=''), width=10, height=10)

  
}
```

## PCA

```{r}
generate_labels <- function(matrix){
labels <- merge(x = as_tibble(rownames(matrix)), y = clinical_data, by.x='value', by.y='ImageNumber', all.x=T)
labels <- labels %>% rename(ImageNumber = value)
# %>%
#   mutate(HER = ifelse(PAM50=='Her2', 'HER2', 'Other')) %>%
#   mutate(LuminalA = ifelse(PAM50=='LumA', 'Luminal A', 'Other')) %>%
#   mutate(LuminalB = ifelse(PAM50=='LumB', 'Luminal B', 'Other')) %>%
#   mutate(Bas = ifelse(PAM50=='Basal', 'Basal', 'Other')) %>%
#   mutate(Norm= ifelse(PAM50=='Normal', 'Normal', 'Other'))

colnames(labels) <- gsub(" ", "_", colnames(labels))
return(labels)
}

plot_pca <- function(matrix, label){
  labels = generate_labels(matrix)
  matrix_withlabels <- cbind.data.frame(matrix, rownames(matrix))
  colnames(matrix_withlabels)[ncol(matrix_withlabels)] = 'ImageNumber'
  matrix_withlabels <- merge(matrix_withlabels, labels, by='ImageNumber', all.x=T)

  res.pca <- prcomp(matrix_withlabels %>% select(-colnames(labels)), scale = FALSE)
  print(fviz_eig(res.pca))
  print(fviz_pca_biplot(res.pca, label="var", habillage=matrix_withlabels %>% pull(!!label), addEllipses = T, ellipse.level = 0.95,select.var = list(contrib = 5) ))
  
}

plot_pca(distance_matrix %>% filter(!(row.names(distance_matrix) %in% getNonTumourImages())), 'PAM50')
ggsave(here('output/Method_comparison/pca/pca_shapeAndScale_PAM50.pdf'),width=8,height=7)

plot_pca(distance_matrixAlternative %>% filter(!(row.names(distance_matrixAlternative) %in% getNonTumourImages())), 'PAM50')
ggsave(here('output/Method_comparison/pca/pca_shapeAndScale_PAM50_alternative.pdf'),width=8,height=7)

plot_pca(distance_matrix %>% filter(!(row.names(distance_matrix) %in% getNonTumourImages())), "ER_HER2_status")
ggsave(here('output/Method_comparison/pca/pca_shapeAndScale_ERHer2.pdf'),width=8,height=7)

plot_pca(distance_matrixAlternative %>% filter(!(row.names(distance_matrixAlternative) %in% getNonTumourImages())),"ER_HER2_status")
ggsave(here('output/Method_comparison/pca/pca_shapeAndScale_ERHer2_alternative.pdf'),width=8,height=7)
```

## UMAP
Plot the parameters as UMAP to see if there are clusters of the same cancer subtype.

```{r}
colnames(generate_labels(shape_matrix))

generate_UMAP <- function(matrix, title){
  labels =generate_labels(matrix)
  umap <- umap(matrix)
  umap <- umap$layout %>%
  as.data.frame()%>%
  rename(UMAP1="V1",
         UMAP2="V2") %>%
  mutate(tnumber=row_number())%>%
  inner_join(labels, by='tnumber')

  return(umap %>%
  ggplot(aes(x = UMAP1, 
             y = UMAP2, 
             color = PAM50))+
  geom_point()+ theme_bw() +
  labs(x = "UMAP1",
       y = "UMAP2",
      subtitle = title))
}

print(generate_UMAP(distance_matrixAlternative,'UMAP plot of shape parameters'))
# print(generate_UMAP(scale_matrix,'UMAP plot of scale parameters'))
```


```

