---
title: "alternative_classification"
author: "Niek Brouwer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(ComplexHeatmap)
library(tidyHeatmap)
library(here)
library(ggplot2)
library(tidyverse)
library(fst)
source(here("UtilityFunctions.R"))
source(here("MEP_UtilityFunctions.R"))
library(patchwork)
library(cowplot)
library(caret)
library(data.table)
library(circlize)


cells <- getCells()
cell_counts <- getCellCounts(cells)
clinical_data <- getClinical()
cell_occurences_combinations <- getCombinationCounts(cells)
all_parameters <- getALLParameters()


```

## Alternative classification
What percentage of the combinations is succesfully estimated for different from and to counts?

```{r}
settings <- tibble(from = seq(1,50),to = seq(1,50)) %>% tidyr::expand(from,to)
for (i in 1:nrow(settings)){
  t <- cell_occurences_combinations %>% filter(n_from >= settings[i,'from'][[1]] & n_to >= settings[i,'to'][[1]])
  settings[i,'percentage'] <- nrow(t %>% filter(unique_sample %in% all_parameters$unique_sample))/nrow(t)*100
}

ggplot(settings,aes(from, to)) + geom_tile(aes(fill=percentage)) + geom_text(aes(label=round(percentage)),size=1) +   scale_fill_gradientn(colours=c("white", "blue"), limits = c(95,100),name='succesful estimation %') + theme_bw()
ggsave(here('output/Distance_results/Missing_estimations_analysis/nbyn_practicalEstimations.pdf'))


```

How many combinations fulfill the min 20 from, 5 to requirement?
```{r}
print(nrow(cell_occurences_combinations %>% filter(n_from >= 20 & n_to >= 5)))
print(nrow(cell_occurences_combinations %>% filter(n_from >= 20 & n_to >= 5))/nrow(cell_occurences_combinations %>% filter(n_from > 0 & n_to > 0))*100)

```
## Alternative classification
What cell types require group lumping?

```{r}
all_possible_samples <- cell_occurences_combinations %>% filter(n_to >= 5 & n_from >= 20) %>% dplyr::count(phenotype_combo) %>%
  separate(phenotype_combo, into=c('phenotype_from','phenotype_to'),sep='_to_', remove = FALSE)

all_possibilities_matrix <- reshape2::dcast(all_possible_samples,phenotype_to  ~ phenotype_from, value.var = 'n')
rownames(all_possibilities_matrix) <- all_possibilities_matrix %>% pull(phenotype_to)
all_possibilities_matrix <- all_possibilities_matrix %>% select(-c(phenotype_to))
all_possibilities_matrix <- all_possibilities_matrix   %>% replace(is.na(.), 0)

library(circlize)
col_fun = colorRamp2(c(0, 50,51, 794), c("black",'black', "white", "red"))

hm1 <- Heatmap(all_possibilities_matrix, cluster_rows = T, cluster_columns = T, col=col_fun, column_names_gp = grid::gpar(fontsize = 6),
  row_names_gp = grid::gpar(fontsize = 6), column_title = 'phenotype_from', row_title = 'phenotype_to')
save_pdf(hm1, filename = here('output/Distance_results/Missing_estimations_analysis/fivetwenty_theoreticalEstimations_HM.pdf'),width=7, height=6)

belowHunderedCount <- all_possible_samples %>% mutate(enoughSamples = ifelse(n > 50, T, F)) %>% group_by(phenotype_from) %>% count(enoughSamples) %>% filter(enoughSamples == FALSE) %>% filter(n > round(length(unique(all_possible_samples$phenotype_from))/2)) %>% pull(phenotype_from)

print(belowHunderedCount)
```
We come up with a new classification based on the expression profiles as shown by Danenberg et al.
We follow the following procedure:
Epithelial cells:
1. Split in CK+ and CK- cells
2. Split in surface marker + and -

For the other cells, we merge groups with unique expression  

```{r}
alternativeClasses = list(# Epithelial
                          # CK+ cells
                          'CK+_ER+_cells'  =  c( "ER^{hi}CXCL12^{+}" , "CK8-18^{+} ER^{hi}"),
                          'HER2+_cells' = c("HER2^{+}"),
                          'CK+_ER-_cells' = c("CK8-18^{hi}ER^{lo}","CK^{+} CXCL12^{+}","CK8-18^{hi}CXCL12^{hi}", "CD15^{+}","Ep CD57^{+}"), 
                          "Basal" = c("Basal"),
                          "CKmed_ERorHER2-_cells" = c("CK^{med}ER^{lo}"),
                          
                          # CK- cells
                          "CK-_cells" = c("CK^{lo}ER^{lo}" ,"CK^{lo}ER^{med}" ),
                          'CK-MHC+_cells' = c("MHC I^{hi}CD57^{+}","MHC I & II^{hi}","MHC^{hi}CD15^{+}"),

                    
                          # Distinct markers
                          'Ki67+_cells_TME' = c("Ki67^{+}"),
                          'Ki67+_cells_epithelial' = c("Ep Ki67^{+}"),
                          'CD4+_cells_APCs' = c('CD4^{+} T cells', 'CD4^{+} T cells & APCs'),
                          'CD8+_cells' = c("CD8^{+} T cells", "T_{Reg} & T_{Ex}"),
                          'CD38+_cells' = c("CD38^{+} lymphocytes"),
                          'CD57+_cells' = c("CD57^{+}"),
                          
                          'Granulocytes_Macrophages'  = c("Granulocytes", "Macrophages & granulocytes", "Macrophages"),
                          'Fibroblasts' = c("Fibroblasts","Fibroblasts FSP1^{+}"),
                          "Myofibroblasts" = c("Myofibroblasts",'Myofibroblasts PDPN^{+}')
                          )

```


Can we improve this number with alternative classification?
```{r}
alternativeClassification <- function(mergelist,removelist){
  names <- names(mergelist)
  for (m in 1:length(mergelist)){
      cell_counts <- cell_counts %>% mutate(meta_description = ifelse(meta_description %in% mergelist[[m]], names[[m]], meta_description))

  }
  
  cell_counts <- cell_counts %>% group_by(tnumber, meta_description) %>% dplyr::summarise(n = sum(n))  %>% 
  ungroup()
  
  cell_counts <- cell_counts %>% filter(!(meta_description %in% removelist))
  
  
  return(cell_counts)
} 


getCombinationCounts <- function(cellcounts){
  cell_counts_to <- cellcounts %>% rename(n_to = n)
  cell_counts_from <- cellcounts %>% rename(n_from = n)
  
  expand_cell_combinations <- cell_counts_to %>% tidyr::expand(tnumber, meta_description, meta_description) %>% 
    rename(phenotype_from = meta_description...2) %>% 
    rename(phenotype_to = meta_description...3)
  
  combination_counts <- merge(expand_cell_combinations, cell_counts_from, by.x=c('tnumber', 'phenotype_from'), by.y=c('tnumber', 'meta_description'))
  combination_counts <- merge(combination_counts, cell_counts_to, by.x= c('tnumber', 'phenotype_to'), by.y=c('tnumber', 'meta_description'))
  combination_counts <- combination_counts %>%
    unite('phenotype_combo', c(phenotype_from, phenotype_to), sep= '_to_', remove=F) %>%
    unite('unique_sample', c(tnumber, phenotype_combo), remove=F)
  
  return(combination_counts)
  
}

testMerge <- function(merge, remove){
  df <- getCombinationCounts(alternativeClassification(merge,remove))
  dfCounts <- df  %>% filter(n_from >= 5 & n_to >= 20) %>% dplyr::count(phenotype_combo) %>%
  separate(phenotype_combo, into=c('phenotype_from','phenotype_to'),sep='_to_', remove = FALSE)
  

  all_possibilities_matrix <- reshape2::dcast(dfCounts,phenotype_to  ~ phenotype_from, value.var = 'n')
  rownames(all_possibilities_matrix) <- all_possibilities_matrix %>% pull(phenotype_to)
  all_possibilities_matrix <- all_possibilities_matrix %>% select(-c(phenotype_to))
  all_possibilities_matrix <- all_possibilities_matrix   %>% replace(is.na(.), 0)
  
  col_fun = colorRamp2(c(0, 794), c("white", "red"))
  

  hm1 <- Heatmap(all_possibilities_matrix, cluster_rows = T, cluster_columns = T, col=col_fun, column_names_gp = grid::gpar(fontsize = 6),
    row_names_gp = grid::gpar(fontsize = 6), column_title = 'phenotype_from', row_title = 'phenotype_to')
  save_pdf(hm1, filename = here('output/Distance_results/Missing_estimations_analysis/ImprovedsignificantSampleCountHM.pdf'),width=7, height=6)
  
  
  belowHunderedCount <- dfCounts %>% mutate(enoughSamples = ifelse(n > 50, T, F)) %>% group_by(phenotype_from) %>% count(enoughSamples) %>% filter(enoughSamples == F) %>% filter(n > round(length(unique(dfCounts$phenotype_from))/2)) %>% pull(phenotype_from)

return(belowHunderedCount)
  
}

print(testMerge(alternativeClasses, list()))


```
```{r}

testMerge <- function(merge,remove){
  df <- getCombinationCounts(alternativeClassification(merge,remove))
  print('percentage of combinations that fulfill 20-5 criteria')
  print(nrow(df %>% filter(n_from >= 20 & n_to >= 20))/nrow(df  %>% filter(n_from > 0 & n_to > 0))*100)
  print('instead of: ')
  print(nrow(cell_occurences_combinations %>%  filter(n_from >= 20 & n_to >= 20))/nrow(cell_occurences_combinations %>% filter(n_from > 0 & n_to > 0))*100)
}

testMerge(alternativeClasses, list())

```
Add new cell type labels

```{r}
changeCellsDF <- function(cells, alternative_classes){
    cells <- cells %>% mutate(original_meta_description = meta_description)

    names <- names(alternative_classes)
    for (m in 1:length(alternative_classes)){
        cells <- cells %>% mutate(meta_description = ifelse(meta_description %in% alternative_classes[[m]], names[[m]], meta_description))
  }
    
  return(cells)
  
}

cells_v2 <- changeCellsDF(getCells(), alternativeClasses)

cells_v2 %>% count(meta_description)


write_fst(cells_v2, here('DATA/SingleCells_altered.fst'))
```
```{r}


```


