---
title: "cohort_subdivision"
author: "Niek Brouwer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Regional subtypes

Both molecular and PAM50 subtypes are diverse in cell type content. This is because of two reasons: cell type stratification is not strict and the classification allows a variety of cells in a class. Secondly, the images are a small sample of the total tumor. The absence of cells doesn't mean that cells are absent in the entire tumor. We therefore divide patient groups further in subtype regions.

```{r}
library(tidyverse)
library(ggplot2)
library(here)
source(here("UtilityFunctions.R"))
source(here("MEP_UtilityFunctions.R"))
library('glmnet', quiet = T)
library('ROCR', quiet = T)
library(ComplexHeatmap)
library(factoextra)
library(corrplot)
library(tidyHeatmap)
library(circlize)

clinical_data <- getClinical()

```
```{r}
generate_labels <- function(matrix){
labels <- merge(x = as_tibble(rownames(matrix)), y = clinical_data, by.x='value', by.y='ImageNumber', all.x=T)
labels <- labels %>% rename(ImageNumber = value)
labels <- labels %>% mutate(ImageNumber = as.integer(labels$ImageNumber))
colnames(labels) <- gsub(" ", "_", colnames(labels))
return(labels)
}

plot_pca <- function(matrix, label){
  labels = generate_labels(matrix)
  matrix_withlabels <- cbind.data.frame(matrix, rownames(matrix))
  colnames(matrix_withlabels)[ncol(matrix_withlabels)] = 'ImageNumber'
  matrix_withlabels <- merge(matrix_withlabels, labels, by='ImageNumber', all.x=T)

  res.pca <- prcomp(matrix_withlabels %>% select(-colnames(labels)), scale = FALSE)
  # print(fviz_eig(res.pca))
  return(fviz_pca_biplot(res.pca, label="var", habillage=matrix_withlabels %>% pull(!!label), addEllipses = T, ellipse.level = 0.95,select.var = list(contrib = 5) ))
  
}

generate_UMAP <- function(matrix){
  labels <- generate_labels(matrix)
  umap <- umap(matrix)
  umap <- umap$layout %>%
  as.data.frame()%>%
  rename(UMAP1="V1",
         UMAP2="V2") %>%
  mutate(ImageNumber=row_number())%>%
  inner_join(labels, by='ImageNumber')

  return(umap %>%
  ggplot(aes(x = UMAP1, 
             y = UMAP2, 
             color = ER_HER2_status))+
  geom_point()+ theme_bw() +
  labs(x = "UMAP1",
       y = "UMAP2",
      subtitle = 'Umap'))
}
```

```{r}
computeEstimationPercentages <- function(matrix){
  na_count <-sapply(matrix, function(y) sum(length(which(is.na(y)))))
  estimation_count <-  1- (na_count / nrow(matrix))
  estimation_count <- data.frame(estimation_count)
  estimations <- as_tibble(data.frame(feature = colnames(matrix),estimation_count))

  estimations <- estimations %>% filter(!(feature %in% c('tnumber', 'PAM50')))
  estimations <- estimations %>% filter(!(feature %in% c(grep('Fibroblasts|Myofibroblasts', estimations$feature,value=T))))
  
  estimations <- estimations %>% arrange(desc(estimation_count)) %>% mutate(rank = row_number()) %>% mutate(rank_percentage = 100* (rank / nrow(estimations)))
  
  return(estimations)
}


```


## Original Classification

```{r}
cellPhenotypes <- getCellProportionsPerImage()
density_matrix <- generate_matrix(cellPhenotypes, 'ImageNumber')

PCAofSubset <- function(type, imageNumbers){
  density_matrix <- density_matrix[row.names(density_matrix) %in% imageNumbers,]
  
  estimations <- computeEstimationPercentages(getShapeFeatures()[row.names(getShapeFeatures()) %in% imageNumbers,])
  dominant_combinations_shape <- estimations %>% filter(rank_percentage < 50) %>% pull(feature)
  dominant_combinations_scale <- estimations %>% filter(rank_percentage < 50) %>% pull(feature)
  shape_features <- getShapeFeatures() %>% select(c('tnumber', dominant_combinations_shape)) %>% rename_with(~paste(., '_shape',sep=''))
  scale_features <- getScaleFeatures() %>% select(c('tnumber', dominant_combinations_scale)) %>% rename_with(~paste(., '_scale',sep=''))
  distance_features <- merge(shape_features,scale_features, by.x='tnumber_shape', by.y='tnumber_scale')
  distance_matrix <- generate_matrix(distance_features,'tnumber_shape')
  distance_matrix <- distance_matrix[row.names(distance_matrix) %in% imageNumbers,]
  
  plot_pca(density_matrix  %>% select(-c('isTumour')) %>% rename_with(~str_remove(., '_CPh')) %>% filter(!(row.names(density_matrix) %in% getNonTumourImages())), 'ER_HER2_status')
  ggsave(paste(here('output/Method_comparison/heatmaps/subtypeRegion_heatmaps/pca_density_'), type, '.pdf',sep=''),width=8,height=7)

  plot_pca(distance_matrix %>% filter(!(row.names(distance_matrix) %in% getNonTumourImages())), "ER_HER2_status")
  ggsave(paste(here('output/Method_comparison/heatmaps/subtypeRegion_heatmaps/pca_shapeAndScale_'), type, '.pdf',sep=''),width=8,height=7)
}

cellCounts <- merge(getDensityFeatures(), clinical_data %>% select(c(ImageNumber, PAM50, ER_HER2_status)), by='ImageNumber')
HER2_images <- cellCounts %>% filter(PAM50 == 'HER2') %>% filter(`HER2^{+}` > 20) %>% pull(ImageNumber)
Basal_images <- cellCounts %>% filter(PAM50 == 'Basal') %>% filter(`Basal` > 20) %>% pull(ImageNumber)
# CD15_images <- cellCounts %>% filter(`CD15^{+}` > 20) %>% pull(ImageNumber)
ERHighCK_images <- cellCounts %>% filter(PAM50 == 'Luminal A') %>% filter(`CK8-18^{+} ER^{hi}` > 20) %>% pull(ImageNumber)
ERHighCX_images <- cellCounts %>% filter(PAM50 == 'LUminal A') %>%  filter(`ER^{hi}CXCL12^{+}` > 20) %>% pull(ImageNumber)


```

## Alternative classification

```{r}
cellCountsAlternative <- getDensityFeaturesAlternative()
CKpos_ERHer2neg_images <- cellCountsAlternative %>% filter(`CK+_ERorHER2-_cells` > 100) %>% pull(ImageNumber)
CKpos_ERHer2pos_images <- cellCountsAlternative %>% filter(`CK+_ERorHER2+_cells` > 100) %>% pull(ImageNumber)
CKmed_ERHer2neg_images <- cellCountsAlternative %>% filter(`CKmed_ERorHER2-_cells` > 100) %>% pull(ImageNumber)
CKneg_images <- cellCountsAlternative %>% filter(`CK-_cells` > 100) %>% pull(ImageNumber)
CKnegMHCpos_images <- cellCountsAlternative %>% filter(`CK-MHC+_cells` > 100) %>% pull(ImageNumber)

cellPhenotypes <- getCellProportionsPerImageAlternative()
density_matrixAlternative <- generate_matrix(cellPhenotypes, 'ImageNumber')

PCAofSubsetAlternative <- function(type, imageNumbers){
  density_matrix <- density_matrixAlternative[row.names(density_matrixAlternative) %in% imageNumbers,]
  
  estimations <- computeEstimationPercentages(getShapeFeaturesAlternative()[row.names(getShapeFeaturesAlternative()) %in% imageNumbers,])
  dominant_combinations_shape <- estimations %>% filter(rank_percentage < 50) %>% pull(feature)
  dominant_combinations_scale <- estimations %>% filter(rank_percentage < 50) %>% pull(feature)
  shape_features <- getShapeFeaturesAlternative() %>% select(c('tnumber', dominant_combinations_shape)) %>% rename_with(~paste(., '_shape',sep=''))
  scale_features <- getScaleFeaturesAlternative() %>% select(c('tnumber', dominant_combinations_scale)) %>% rename_with(~paste(., '_scale',sep=''))
  distance_features <- merge(shape_features,scale_features, by.x='tnumber_shape', by.y='tnumber_scale')
  distance_matrix <- generate_matrix(distance_features,'tnumber_shape')
  distance_matrix <- distance_matrix[row.names(distance_matrix) %in% imageNumbers,]
  
  plot_pca(density_matrix  %>% select(-c('isTumour')) %>% rename_with(~str_remove(., '_CPh')) %>% filter(!(row.names(density_matrix) %in% getNonTumourImages())), 'PAM50')
  ggsave(paste(here('output/Method_comparison/heatmaps/subtypeRegion_heatmaps/pca_density_'), type, '_alternative.pdf',sep=''),width=8,height=7)

  plot_pca(distance_matrix %>% filter(!(row.names(distance_matrix) %in% getNonTumourImages())), "PAM50")
  ggsave(paste(here('output/Method_comparison/heatmaps/subtypeRegion_heatmaps/pca_shapeAndScale_'), type, '_alternative.pdf',sep=''),width=8,height=7)
}

PCAofSubsetAlternative('CKpos_ERHer2neg', CKpos_ERHer2neg_images)
# PCAofSubsetAlternative('CKpos_ERHer2pos', CKpos_ERHer2pos_images)
# PCAofSubsetAlternative('CKmeds_ERHer2neg', CKmed_ERHer2neg_images)
# PCAofSubsetAlternative('CKneg', CKneg_images)
# PCAofSubsetAlternative('CKnegMHCpos', CKnegMHCpos_images)


```


