---
title: "region_analysis"
author: "Niek Brouwer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Region analysis
In this document we are going to investigate the following hypothesis:

What are common spatial arrangements in structure-dense slide regions such as vascular stroma, granulocyte enriched, APC enriched, and suppressed expansion?
We first look into these regions because they are associated with an improved (granulocyte enriched, APC enriched, and suppressed expansion) and decreased (vascular stroma) survival in ER+ patients.

```{r,include = FALSE}
library(tidyverse)
library(ggplot2)
library(here)
library(fst)
library(data.table)
source(here("UtilityFunctions.R"))
source(here('MEP_UtilityFunctions.R'))
source(here('figure_styling.R'))

library(ComplexHeatmap)
library(spatstat)
library(dbscan)
library(patchwork)
library(cowplot)
library(ggpubr)
library(rstatix)
library(rlang)

outdir <- here('output/final_figures/regions/')

clinical_data <- getClinical()
cells <- getCellsAlternative()
structures <- getStructures()
structureLabels <- gsub("[^A-Za-z0-9+[:space:]]",'', structures$TME$labels)
structureLabels <- tibble(index = seq(1,10), label = structureLabels)
TMEStructures <- read_fst(here('scratch/TMEStructures.fst'), as.data.table=T)
allCommunities <- read_rds(here('scratch/allCommunities.rds'))

# structure_type <- 'APC enriched'
# structure_type <- 'Suppressed expansion'
# structure_type <- 'Granulocyte enriched'
structure_type <- "TLSlike"
# structure_type <- "FSP1+ enriched"
# structure_type <- "PDPN+ active stroma"
# structure_type <- "CD8+ macrophages"
# structure_type <- "Active IR"
# structure_type <- "Active stroma"
# structure_type <- 'Vascular stroma'
# 


```

```{r}
allCommunities <- merge(allCommunities, TMEStructures %>% dplyr::select(communityUID, TMEStructure), by='communityUID')

cellsWithoutExpression <- cbind(cells[,1:(ncol(cells)-43)],cells[,(ncol(cells)-3):ncol(cells)])
cellsWithStructure <- merge(cellsWithoutExpression, allCommunities %>% dplyr::select(-c('to', 'to_meta_id', 'nInteractions')), by.x = c('ImageNumber', 'ObjectNumber'), by.y=c('ImageNumber', 'from'), all=T)
cellsWithStructure <- cellsWithStructure %>% mutate(TMEStructure = as.numeric(TMEStructure))
cellsWithStructure <- merge(cellsWithStructure, structureLabels, by.x='TMEStructure', by.y='index',all.x=T) %>% distinct()

```

## KDE separation

Split regions in  images

```{r}
data_with_assigned_locations <- readRDS(paste(here('scratch/Regions_'),gsub(' ','_',gsub('+', 'plus',structure_type,fixed=T),fixed=T), '.rds', sep=''))
# data_with_assigned_locations <- data_with_assigned_locations %>% mutate(assigned_loc = ifelse(assigned_loc == 'S', structure_type, 'low-density region'))
# data_with_assigned_locations <- data_with_assigned_locations %>% mutate(correct_prediction = ifelse(assigned_loc == label, T, F))

data_with_assigned_locations$TMEStructure <- as.character(data_with_assigned_locations$TMEStructure)

show_slideStructures <- function(df, sample_number, column){
  slide_1 <- ggplot(data=df%>% filter(ImageNumber == sample_number)) + geom_point(aes(x=Location_Center_X, y=Location_Center_Y, color=get(column)), alpha =1) +theme_bw() + ggtitle(paste('imagenumber: ', sample_number)) + theme(legend.position = "none") + xlab('') + ylab('') +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank()  #remove y axis ticks
        ) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_color_manual(values = c("#888888","white"))

  return(slide_1)
}


show_slideStructures(data_with_assigned_locations, 3, 'label')
ggsave(here('output/PresentationFigures/slideWithStructures1.pdf'),width =4, height=4)
show_slideStructures(data_with_assigned_locations, 3, 'label')
ggsave(here('output/PresentationFigures/slideWithStructures2.pdf'),width =4, height=4)
show_slideStructures(data_with_assigned_locations, 3, 'TMEStructure')
ggsave(here('output/PresentationFigures/slideWithStructuresALL.pdf'),width =4, height=4)


# for (i in images){
#   print(plot_grid(show_slideStructures(data_with_assigned_locations,i,'label'),show_slideStructures(data_with_assigned_locations,i,'assigned_loc')))
#
# }
#
# splittedCells <- data_with_assigned_locations %>% unite('Split_ImageNumber',c(ImageNumber, assigned_loc), sep = '_', remove = F)
# saveRDS(splittedCells, paste(here('DATA/splittedCells'),'_',gsub(' ', '_',gsub('+', 'plus',structure_type,fixed=T),fixed=T),'.rds', sep=''))


```

## Spatial relationship analysis


```{r}
# structure_type <- gsub(' ', '_', structure_type, fixed=T)
# 
# 
# source(here('loadData.R'))
# 
# CompileALLParameters_regions <- function(type, save=TRUE){
#   all_parameters <- get_parameters(paste(here('scratch/success_models_'),type,'/',sep='')) %>%
#     separate(phenotype_combo, into=c('phenotype_from','phenotype_to'),sep='_to_', remove = FALSE) %>%
#     filter(a > 0.5) %>%
#     filter(b > 8) %>%
#     dplyr::rename(shape =a , scale = b) %>%
#     unite('unique_sample', c(tnumber, phenotype_combo), remove=F)
# 
#   if(save){
#     saveRDS(all_parameters, file = paste(here('scratch/all_parameters_'), type,'.rds',sep=''))
#   }
# 
#   return(all_parameters)
# 
# }
# 
# CompileALLParameters_regions(gsub('+', 'plus',structure_type,fixed=T))
# transform_parameters_to_matrix(readRDS(paste(here('scratch/all_parameters_'), gsub('+', 'plus',structure_type,fixed=T), '.rds',sep='')),
#                                paste(here('scratch/features/regions/shape_parameters_'),gsub('+', 'plus',structure_type,fixed=T), '.rds',sep=''),
#                                paste(here('scratch/features/regions/scale_parameters_'), gsub('+', 'plus',structure_type,fixed=T), '.rds',sep='') )
```

```{r}
structure_type <- gsub(' ', '_', structure_type, fixed=T)


allparameters_structure <- readRDS(paste(here('scratch/all_parameters_'), gsub('+', 'plus',structure_type,fixed=T), '.rds',sep='')) %>% separate(tnumber,c('tnumber','region'),sep='_') %>% mutate(region = ifelse((region == 'low' | region == 'low-density region'),'other',paste(gsub('_', ' ', structure_type),'')))  %>% filter(!(tnumber %in% getNonTumourImages()))


tumor_celltypes <- grep('CK', unique(cells %>% pull(meta_description)),fixed = T,value=T)
tme_celltypes <- setdiff( unique(cells %>% pull(meta_description)), tumor_celltypes)

ggplot() + 
  geom_point(data = allparameters_structure, aes(x=shape, y=scale,colour=region),alpha=0.8, size=0.5) +
  theme_bw() +
  ylim(10,300) + xlim(0,6) +
  scale_y_log10() +
  xlab('Shape') + ylab('Scale') + ggtitle("Parameters of structure regions") + 
  guides(colour = guide_legend(override.aes = list(size=10)))

```

Are some features estimated more in a certain region?

```{r}
shapeFeatures_structure <- readRDS(paste(here('scratch/features/regions/shape_parameters_'),gsub('+', 'plus',structure_type,fixed=T), '.rds',sep=''))%>% separate(tnumber,c('tnumber','region'),sep='_') %>% mutate(region = ifelse((region == 'low' | region == 'low-density region'),'other',paste(gsub('_', ' ', structure_type),'')))
scaleFeatures_structure <- readRDS(paste(here('scratch/features/regions/scale_parameters_'), gsub('+', 'plus',structure_type,fixed=T), '.rds',sep='') ) %>% separate(tnumber,c('tnumber','region'),sep='_') %>% mutate(region = ifelse((region == 'low' | region == 'low-density region'),'other',paste(gsub('_', ' ', structure_type),'')))

na_counts <- merge(data.frame(colSums(is.na(shapeFeatures_structure %>% filter(region == paste(gsub('_', ' ', structure_type),''))))),data.frame(colSums(is.na(shapeFeatures_structure %>% filter(region == 'other')))),by='row.names')

colnames(na_counts)[1] <- "feature"
colnames(na_counts)[2] <- "high"
colnames(na_counts)[3] <- "low"
na_counts <- reshape2::melt(na_counts)
na_counts <- na_counts %>% filter(!(feature %in% c('tnumber', 'region'))) %>% separate(feature, c('phenotype_from', 'phenotype_to'),sep='_to_') 
# na_counts <- na_counts %>% mutate(value = 100*value/length(unique(shapeFeatures_structure$tnumber)))

ggplot(na_counts %>% filter(variable=='high') %>% filter((phenotype_from %in% tme_celltypes) & (phenotype_to %in% tme_celltypes))) + geom_tile(aes(x=phenotype_from, y=phenotype_to, fill=value)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle('high density region')

ggplot(na_counts %>% filter(variable=='low' | variable == 'low-density region') %>% filter((phenotype_from %in% tme_celltypes) & (phenotype_to %in% tme_celltypes))) + geom_tile(aes(x=phenotype_from, y=phenotype_to, fill=value)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle('low density region')

```

```{r}
generate_matrix_zeroOneScaling <- function(df, col_rownames ,subselection=NULL, NA_percentage=0, scale_rows=FALSE){
  m <- as.data.frame(df)
  rownames(m) <- m[,col_rownames]
  m <- subset(m,select = -c(get(col_rownames)))
  
  # Filtering
  m <- m[, which(colMeans(!is.na(m)) > NA_percentage)]

  m <- replace(m,is.na(m),0)
  m <- as.data.frame(m)
  
  
  # Scaling
  # m <- apply(m, 2, function(x) (x - min(x)) / (max(x) - min(x)))
  
  
  return(m)
}

shapeFeatures_structure <- generate_matrix_zeroOneScaling(readRDS(paste(here('scratch/features/regions/shape_parameters_'),gsub('+', 'plus',structure_type,fixed=T), '.rds',sep='')), 'tnumber')
scaleFeatures_structure <- generate_matrix_zeroOneScaling(readRDS(paste(here('scratch/features/regions/scale_parameters_'), gsub('+', 'plus',structure_type,fixed=T), '.rds',sep='')), 'tnumber')
densityFeatures_structure <- generate_matrix_zeroOneScaling(readRDS(paste(here('scratch/features/regions/cell_proportions_per_image_'), gsub('+', 'plus',structure_type,fixed=T),'.rds',sep='')), 'ImageNumber')

filterSignificant <- function(df){
  subset <- as.data.frame(df) %>% mutate(tnumber = row.names(df)) %>% separate(tnumber, c('tnumber', 'group'),sep = '_')
  nonZero_counts <- subset %>%
    group_by(group) %>%
    summarise(across(everything(), ~ sum(. != 0)))
  melted_nonzero <- as.data.frame(t(nonZero_counts %>% dplyr::select(-c('group','tnumber'))))
  melted_nonzero <- melted_nonzero %>% mutate(sig = ifelse((as.numeric(V1) > 5) &(as.numeric(V2) > 5), T, F))
  
  significant_features <- rownames(data.frame(melted_nonzero[2:nrow(melted_nonzero),]) %>% filter(sig == T))
  
  return(significant_features)

}



shapeFeatures_structure <- shapeFeatures_structure[,filterSignificant(shapeFeatures_structure)] + 0.001
scaleFeatures_structure <- scaleFeatures_structure[,filterSignificant(scaleFeatures_structure)] + 0.001

```

```{r}
# Load required libraries
library(multcomp)
library(ggrepel)
fold_threshold = 1
p_value = -log10(0.05)

computePvalues <- function(featureSet){
  subset <- as.data.frame(featureSet) %>% mutate(tnumber = row.names(featureSet)) %>% separate(tnumber, c('tnumber', 'group'),sep = '_') %>% dplyr::select(-dplyr::any_of(c('tnumber'))) %>% mutate(group = ifelse((group == 'low' | group == 'low-density region'),'other',paste(gsub('_', ' ', structure_type),'')))%>% dplyr::select(-any_of(c('isTumour')))
  
  
  # Create an empty vector to store the adjusted p-values and log2 fold changes
  p_values <- rep(NA, ncol(subset)-1)
  log2_fold_changes <- rep(NA, ncol(subset)-1)

  # Perform univariate feature importance analysis for each feature
  for (i in 1:(ncol(subset)-1)) {
    # Extract the current feature
    
    feature_subset <- subset %>% filter(!(subset[, i] == 0.001 ))
    # feature_subset <- subset

    feature <- feature_subset[, i]
    response <- feature_subset$group
    
    computePvalue <- function(feature, response) {
    p_value <- tryCatch(
        { p <- t.test(feature ~ response)$p.value
        return(p)
        
        
        },
        error=function(cond) {
            message(colnames(featureSet)[[i]])
            return(1)
        })
        return(p_value)
    }
    
    p_value <- computePvalue(feature, response)


    # Calculate the log2 fold change between the response groups
    mean_response <- tapply(feature, response, mean)

    log2_fold_change <- log2((mean_response[paste(gsub('_', ' ', structure_type),'')]) / mean_response['other'])

    # Store the p-value and log2 fold change in their respective vectors
    p_values[i] <- p_value
    log2_fold_changes[i] <- log2_fold_change
  }

  # Apply p-value adjustment method (e.g., Bonferroni, Benjamini-Hochberg)
  adjusted_p_values <- p.adjust(p_values, method = "BH")

  # Create a data frame with the features, p-values, log2 fold changes, and adjusted p-values
  result <- data.frame(Feature = colnames(subset %>% dplyr::select(-c(group))),
                       p_value = p_values,
                       log2FoldChange = log2_fold_changes,
                       Adjusted_p_value = adjusted_p_values,
                        x_dense = mean_response[paste(gsub('_', ' ', structure_type),'')],
                        x_sparse = mean_response['other'])%>%
            mutate(Sig = ifelse(Adjusted_p_value < 0.05, T, F))

  return(result)
}


univariateFeatures <- computePvalues(shapeFeatures_structure) %>% mutate(featureSet = 'shape')
univariateFeatures <- rbind(univariateFeatures, (computePvalues(scaleFeatures_structure) %>% mutate(featureSet = 'scale')))
univariateFeatures <- rbind(univariateFeatures, (computePvalues(densityFeatures_structure) %>% mutate(featureSet = 'density')))
univariateFeatures <- univariateFeatures %>% mutate(minlogP = -log(Adjusted_p_value, base = 10))
univariateFeatures <- univariateFeatures %>% mutate(Feature = gsub('_CPh','', Feature,fixed=T))

# univariateFeatures <- univariateFeatures %>% mutate(Sig = ifelse((minlogP > -log10(0.05)) & ((abs(log2FoldChange) > 1)),T,F))
univariateFeatures <- univariateFeatures %>% mutate(Sig = ifelse((minlogP > -log10(0.05)),T,F))

univariateFeatures <- univariateFeatures %>% filter(Feature %in% grep('CK', univariateFeatures$Feature, invert = T, fixed = T,value = T))

```

```{r}
# Creating the volcano plot

univariateFeatures <- univariateFeatures %>% mutate(`features` = ifelse(featureSet %in% c('shape', 'scale'), 'spatial feature', ifelse(featureSet == 'original_densities', 'fraction \n Danenenberg classification', 'fraction')))

univariateFeatures <- univariateFeatures %>% filter(!(Feature == "Myofibroblasts_to_Fibroblasts"))

top_values1 <- rbind(univariateFeatures %>% filter(Sig == T) %>% filter(log2FoldChange < 0) %>% filter(features == 'spatial feature') %>% slice_max(order_by = (minlogP), n = 5),univariateFeatures %>% filter(Sig == T)%>% filter(log2FoldChange < 0) %>% filter(features == 'fraction')) %>% distinct()
top_values2 <- rbind(univariateFeatures %>% filter(Sig == T) %>% filter(log2FoldChange > 0) %>% filter(features == 'spatial feature') %>% slice_max(order_by = (minlogP), n = 5),univariateFeatures %>% filter(Sig == T)%>% filter(log2FoldChange > 0) %>% filter(features == 'fraction')) %>% distinct()

top_values2 <- top_values2 %>% mutate(Feature = ifelse(Feature == 'Granulocytes_Macrophages_to_Granulocytes_Macrophages','Granulocytes_Macrophages_to\n_Granulocytes_Macrophages', Feature))


p1 <- ggplot() +
  # geom_vline(xintercept=c(-fold_threshold, fold_threshold), col="black") +
  geom_hline(yintercept=-log10(0.05), col="black") +
  geom_vline(xintercept = 0,linetype=1,color='black',alpha=0.5) +
  geom_point(data = (univariateFeatures %>% filter(Sig == T)), aes(x = log2FoldChange, y =minlogP,colour=features),alpha=0.8) +
  geom_point(data = (univariateFeatures %>% filter(Sig == F)), aes(x = log2FoldChange, y =minlogP),color='grey',alpha=0.8) +
  geom_label_repel(data= univariateFeatures %>% filter(Sig == T), aes(x=log2FoldChange, y=minlogP, label=gsub('or','',gsub('_',' ',gsub('_originalType','',Feature))),fill=features),segment.color='black',min.segment.length = 0, color='white',size =4, ylim=c(2,NA),force=50) +
    # geom_label_repel(data= top_values1, aes(x=log2FoldChange, y=minlogP, label=gsub('or','',gsub('_',' ',gsub('_originalType','',Feature))),fill=features),segment.color='black',min.segment.length = 0, color='white',size =4, xlim=c(NA,-2),ylim=c(2,NA),force=50) +
    # geom_label_repel(data= top_values2, aes(x=log2FoldChange, y=minlogP, label=gsub('or','',gsub('_',' ',gsub('_originalType','',Feature))),fill=features),segment.color='black',min.segment.length = 0, color='white',size =4, xlim=c(1.5,NA),ylim=c(2,NA),force=200) +
  
  labs(x = "Log2 Fold Change",
       y = "-log10 Adjusted P-value") +
  theme_bw() + scale_x_continuous(limits = c(-5, 5)) + ggtitle(gsub('_', ' ', structure_type)) + theme(plot.title = element_text(hjust=0.5)) + scale_color_manual(breaks = c('density','shape','scale'),values = getDiscreteColors(4)[1:3]) + scale_colour_manual(name = '', breaks = c('fraction','spatial feature'),values = getDiscreteColors(3)[1:2])  + scale_fill_manual(name = '', breaks = c('fraction','spatial feature'),values = getDiscreteColors(3)[1:2]) + guides(colour = FALSE)


ggsave(paste(outdir,'volcanoPlot_', gsub('+', 'plus',structure_type,fixed=T), '.pdf',sep=''),width=6,height=4)

```



```{r}
significant_densityFeatures <- grep('_to_', univariateFeatures %>% filter(Sig == T) %>% pull(Feature),invert=T,value=T, fixed=T)
significant_distanceFeatures <- unique(grep('_to_',univariateFeatures %>% filter(Sig == T) %>% pull(Feature),value = T))


print(significant_densityFeatures)
print(significant_distanceFeatures)

# print(grep(paste(significant_densityFeatures,collapse="|"), significant_distanceFeatures,invert=T,value=T))
# 
grep(paste(c('B cells', 'CD8', 'CD4','Granulocytes_Macrophages', 'Fibroblasts', 'Myofibroblasts'),collapse="|"), significant_distanceFeatures,value=T, invert =T)


#Vascular stroma
# significant_distanceFeatures <- c('Granulocytes_Macrophages_to_Fibroblasts')

#Active stroma
# significant_distanceFeatures <- c("Fibroblasts_to_Fibroblasts","Myofibroblasts_to_Fibroblasts", "CD8+_cells_to_B cells")

significant_distanceFeatures <- c("Myofibroblasts_to_CD8+_cells", "Myofibroblasts_to_B cells")


```


```{r}
retrieveDistanceData <- function(type, parameter_df){
  distance_data <- read_tsv(paste(here('scratch/AUCScaledSlides_300_'), type, '.tsv',sep='')) %>% as_tibble %>%
  mutate(distance_window = WinMean) %>%
  mutate(phenotype_combo = paste(phenotype_from, phenotype_to, sep='_to_')) %>%
  dplyr::select(tnumber, phenotype_combo, `N.per.mm2.scaled`, distance_window) %>%
  mutate(new = `N.per.mm2.scaled` * 1000) %>%
  mutate(new =round(new))
  
  estimated_distances <- distance_data  %>% unite('unique_sample', c(tnumber, phenotype_combo), remove=F) %>% filter(unique_sample %in% parameter_df$unique_sample)
  
  return(estimated_distances)
  
}

computeMedians <- function(distance_data){
    total_observations <- distance_data %>% group_by(unique_sample) %>%
  summarise(total_observations= c_across(cols = c(N.per.mm2.scaled)) %>% sum())

  medians <- merge(distance_data, total_observations, by='unique_sample', all.x=T)
  medians <- medians %>% mutate(frequency = round((N.per.mm2.scaled / total_observations)*10000))
  medians <- medians %>% group_by(unique_sample) %>% summarise(
    median_observations = median(rep(distance_window,frequency))
    )
  return(medians)
}
```



```{r}
densities <- readRDS(paste(here('scratch/features/regions/cell_proportions_per_image_'),gsub('+', 'plus',structure_type,fixed=T),'.rds',sep = ''))
names(densities) <- gsub('_CPh','',gsub('+', 'plus', names(densities), fixed = T))
names(densities) <- gsub('_CPh','',gsub('-', 'min', names(densities), fixed = T))

distance_data_structure <- retrieveDistanceData(gsub('+', 'plus',structure_type,fixed=T), allparameters_structure)
medians_structure <- computeMedians(distance_data_structure)

colors <- tibble(name = c(paste(gsub('_', ' ', structure_type),''),'other'), color = getDiscreteColors(3)[1:2])

```

```{r}

DensityPlot <- function(df, response, feature){
  d <- df %>% separate(ImageNumber, c('tnumber', 'region'),sep='_') %>% mutate(region = ifelse((region == 'low' | region == 'low-density region'),'other',paste(gsub('_', ' ', structure_type),'')))
  
  p <- ggboxplot(na.omit(d), x = 'region', y = feature, fill='region') + ylab('cell type fraction') + xlab('') + ggtitle(gsub('plus', '+', gsub('min', '-', feature, fixed = T), fixed = T)) + stat_compare_means(method = "t.test", label.x.npc = 'center', label.y.npc = 'top',label = 'p.format',color='red') + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + scale_fill_manual(breaks = c(paste(gsub('_', ' ', structure_type),''),'other' ), values=getDiscreteColors(3)[1:2]) + theme(legend.position = "none") + theme(plot.title = element_text(size = 10,hjust=0.5))
  return(p)
}

DensityPlot(densities, structure_type, 'B cells')
ggsave(paste(outdir, 'TLS_like_B_cells.pdf',sep=''),width=5,height=4)
```



```{r}

retrieveParameters <- function(df,structure,feature){
  d <- data.frame(df) %>% dplyr::select(-dplyr::any_of(c('tnumber'))) %>% filter(phenotype_combo == feature) %>% rename(group = region)

  mean_d <- d %>% group_by(group) %>% summarise(mean_shape = mean(shape), mean_scale =mean(scale))
  # print(mean_d)
  shape_group <- mean_d %>% filter(group == paste(gsub('_', ' ', structure_type),'')) %>% pull(mean_shape)
  scale_group <-  mean_d %>% filter(group == paste(gsub('_', ' ', structure_type),'')) %>% pull(mean_scale)
  shape_nogroup <- mean_d %>% filter(group == 'other') %>% pull(mean_shape)
  scale_nogroup <- mean_d %>% filter(group == 'other') %>% pull(mean_scale)
  
  return(list('shape' = shape_group, 'scale' = scale_group, 'shape_alternative' = shape_nogroup, 'scale_alternative' = scale_nogroup))
}


create_boxplot <- function(df,medians, structure,feature){
  d <- data.frame(df) %>% mutate(tnumber = row.names(df)) %>% dplyr::select(-dplyr::any_of(c('tnumber')))%>% filter(phenotype_combo == feature) %>% rename(group = region) %>% mutate(group = ifelse(grepl(structure, unique_sample,fixed =T), paste(gsub('_', ' ', structure_type),''), 'other'))
  
  medians <- medians %>% filter(grepl(feature, unique_sample,fixed = T) ) %>% mutate(group = ifelse(grepl(structure, unique_sample,fixed =T), paste(gsub('_', ' ', structure_type),''), 'other'))
  
    d$group <- factor(d$group, levels = c( paste(gsub('_', ' ', structure_type),''), 'other'))
  medians$group <- factor(medians$group, levels = c( paste(gsub('_', ' ', structure_type),''), 'other'))
  
  b1 <- ggboxplot(na.omit(d),x='group', y='shape',fill='group') + theme_bw() + xlab('') + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(), axis.ticks.x=element_blank(),legend.position = "none") + scale_fill_manual(breaks = c(paste(gsub('_', ' ', structure_type),''),'other' ), values=getDiscreteColors(3)[1:2]) + stat_compare_means(method = "t.test", label = "p.format", label.y.npc =0.9, label.x.npc = 'left',size=3,color='red') + plot_annotation(tag_level='A')
  b2 <- ggboxplot(na.omit(d), x='group', y='scale',fill='group') + theme_bw() + xlab('') + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),legend.position = "none") + scale_fill_manual(breaks = c(paste(gsub('_', ' ', structure_type),''),'other' ), values=getDiscreteColors(3)[1:2])  + stat_compare_means(method = "t.test", label = "p.format", label.y.npc =0.9, label.x.npc = 'left',size=3,color='red') + plot_annotation(tag_level='A')
  b3 <- ggboxplot(medians, x='group', y='median_observations',fill='group') + theme_bw() + ylab('median') + xlab('') + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),legend.position = "none") + scale_fill_manual(breaks = c(paste(gsub('_', ' ', structure_type),''),'other' ), values=getDiscreteColors(3)[1:2])  + stat_compare_means(method = "t.test", label = "p.format", label.y.npc =0.9, label.x.npc = 'left',size=3,color='red') + plot_annotation(tag_level='A')
  
  return(b1 + b2 + b3)
  
}

showDistanceCurves <- function(df,medians, structure, feature){
    params <- retrieveParameters(df,structure,feature)
    print(params)
    g1 <- ggplot() + 
      stat_function(fun = dweibull, args = list(shape = params$shape,scale = params$scale), aes(colour=paste(gsub('_', ' ', structure_type),'')),n=600)  +
      stat_function(fun = dweibull, args = list(shape = params$shape_alternative ,scale = params$scale_alternative), aes(colour='other'),n=600)  +
    xlim(0,600) + theme_bw() + xlab('1-NN distance (micron)') + ylab('normalized counts') + ggtitle(gsub('_', ' ',feature,fixed=T)) + scale_color_manual(name='region',breaks = c(paste(gsub('_', ' ', structure_type),''),'other' ), values=getDiscreteColors(3)[1:2])  + theme(plot.title = element_text(size = 10,hjust=0.5))  + theme(legend.position=c(.85,.23))
      
    b1 <- create_boxplot(df,medians, structure,feature)
    g_total <- g1 + inset_element(b1, 0.3, 0.5, 0.98, 0.98)

    return(g_total)

}


# Show examples for every subtype
for (i in significant_distanceFeatures){
  p1 <- DensityPlot(densities,gsub('_', ' ',structure_type),gsub('+','plus',gsub('-','min',str_split(i,'_to_')[[1]][[1]], fixed=T),fixed=T)) +DensityPlot(densities,gsub('_', '',structure_type),gsub('+','plus',gsub('-','min',str_split(i,'_to_')[[1]][[2]],fixed=T),fixed=T)) + showDistanceCurves(allparameters_structure,medians_structure, gsub('_',' ',structure_type), i) +   plot_layout(widths = c(1, 1, 3 )) + plot_annotation(tag_level='A')


  ggsave(paste(outdir,'plot_',gsub('+', 'plus',structure_type,fixed=T),gsub('+','plus',i,fixed=T),'.pdf', sep=''),width=14,height=7)

}

```

Special plot for TLS-like neighborhood.
```{r}
DensityPlot(densities,gsub('_', ' ',structure_type),'Myofibroblasts') + showDistanceCurves(allparameters_structure,medians_structure, gsub('_',' ',structure_type), 'CD8+_cells_to_Myofibroblasts') +DensityPlot(densities,gsub('_', '',structure_type),'CD8plus_cells') + showDistanceCurves(allparameters_structure,medians_structure, gsub('_',' ',structure_type), 'Myofibroblasts_to_CD8+_cells')  + plot_layout(ncol=2,widths = c(1, 3, 1,3)) + plot_annotation(tag_level='A')

ggsave(paste(outdir,'plot_',gsub('+', 'plus',structure_type,fixed=T),gsub('+','plus','Myofibroblasts_to_CD8+_cells',fixed=T),'.pdf', sep=''),width=12,height=14)

```

